---
title: "SPE data Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

R markdown notebook for universal data processing of the SPE work. All of the chunks have explanatory text. Please take care to read this information as you work through the notebook.



# Project set-up

To set up the project please use the following template

\~[path]/[projectID]/ - *IMPORTANT NOTE - THIS WILL NOT WORK IF THE PATH LENGTH IS > 260 CHARACTERS (WINDOWS)*

------------------------------------------------------------------------

\~[path]/[projectID]/AB/ \~[path]/[projectID]/AB/data/mzML - *(copy all AB mzML to this directory)*

------------------------------------------------------------------------

\~[path]/[projectID]/N/

\~[path]/[projectID]/N/data/mzML - *(copy all N mzML to this directory)*

------------------------------------------------------------------------

\~[path]/[projectID]/msMethods - *(copy both the AB and N ms method (.exp) directly from the waters instrument into this directory)*

------------------------------------------------------------------------



## Load packages and set-up master list

------------------------------------------------------------------------

*USER INPUT REQUIRED*

------------------------------------------------------------------------


```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#load packages
package_list <- c('statTarget', 'svDialogs', 'ggpubr', 'janitor', 'mzR', 'cowplot', 'tidyverse', 'fedmatch','sva', 'gghalves')

for(idx_package in package_list){
  if(length(which(row.names(installed.packages()) == idx_package)) > 0){
    suppressMessages(require(package = idx_package,
                             character.only = TRUE))
  } else {
    svDialogs::dlg_message(
      paste0("the package ", idx_package, " is not installed. Please install ", idx_package, " before continuing.")
    )
  }
}

```


```{r}

#set up project master list
master_list <- list(); master_list$environment <- list(); master_list$environment$user_functions <- list(); master_list$templates <- list(); master_list$templates$mrm_guides <- list(); master_list$project_details <- list();    master_list$data <- list(); master_list$data$mzR <- list(); master_list$summary_tables <- list(); master_list$process_lists <- list(); master_list$plots <- list(); master_list$plots$chromatogram <- list()

#store environment details
master_list$environment$r_version <- sessionInfo()$R.version$version.string
master_list$environment$base_packages <- sessionInfo()$basePkgs
master_list$environment$user_packages <- paste0(names(sessionInfo()$otherPkgs), ": ", paste0(installed.packages()[names(sessionInfo()$otherPkgs), "Version"]))

#######USER INPUT#############

#create project folder structure and normalise path
svDialogs::dlg_message("select project directory [~PATH/PROJECT_ID/"); master_list$project_details$project_dir <-  rstudioapi::selectDirectory(); master_list$project_details$project_dir <- normalizePath(master_list$project_details$project_dir, winslash = "/")
#set SPE type
master_list$project_details$spe <- svDialogs::dlg_input("is this project MCX or MAX?", "MCX/MAX")$res
#set wash type
master_list$project_details$wash <- svDialogs::dlg_input("is this project AB or N?", "AB/N")$res
#set user
master_list$project_details$user_name <- svDialogs::dlg_input("user initials: ", "")$res
#set project name
master_list$project_details$project_name <-  svDialogs::dlg_input("project name: ", "")$res

#set filePath
filePath <- paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash,"/data/mzml")

```



## use r to organise mzml into subplates

```{r}

svDialogs::dlg_message(paste0("set up mzml folder in project directory as follows:  " , gsub(".*/", "~/", master_list$project_details$project_dir), "/", master_list$project_details$wash, "/data/mzml")); 
svDialogs::dlg_message("copy all mzmls into this folder, do not sort into subfolders");
svDialogs::dlg_message("NOTE THIS SCRIPT WILL NOT WORK ON WINDOWS IF PATH LENGTH IS >260 CHARACTERS")
#set filePath
filePath <- paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash,"/data/mzml")

#first move mzML into individual plates
fileList <- list.files(filePath, pattern = ".mzML", recursive = FALSE, full.names = TRUE)
svDialogs::dlg_message(paste0("there are ", length(fileList), " mzml files"))

if(length(fileList) > 0){

pathLength = nchar(fileList[1])
svDialogs::dlg_message(paste0("Your path length is: ", pathLength))


if(pathLength >= 240){
  svDialogs::dlg_message(paste0("Your path length is approaching the Windows maximum of 260. If you are on Windows it will fail. Move project to a shorter length directory"))
}

if(pathLength < 240){
  svDialogs::dlg_message(paste0("Your path length looks OK and is shorter than the Windows maximum of 260."))
}

#only perform the following steps if there are mzMLs in the folder
if(length(fileList) > 0 & pathLength < 240){
  
#sort files into plate subfolders
plateList <-  fileList %>% gsub(".*SS", "SS", .) %>% gsub(".*p0", "p0", .) %>% gsub("_.*", "", .) %>% unique()

idxPlate = plateList[1]
for(idxPlate in plateList){
  plateMZML <- fileList[grepl(idxPlate, fileList)]

  if(!dir.exists(paste0(filePath, "/", idxPlate))){
    dir.create(paste0(filePath, "/", idxPlate))
  }#if loop

  idxMZML = plateMZML[1]
  for(idxMZML in plateMZML){

  file.rename(
    from = idxMZML,
    to = sub("data/mzml/", paste0("data/mzml/", idxPlate, "/"), idxMZML)
  )

  } #for idxMZML

}#for idxPlate

} #if length mzML > 0

updatedFileList <- list.files(filePath, pattern = ".mzML", recursive = TRUE, full.names = TRUE); updatedFileList <- updatedFileList[!grepl("archive",updatedFileList)]; #length(updatedFileList)
plateList <- gsub(".*p0", "p0", updatedFileList) %>% gsub("_.*", "", .) %>% unique()

svDialogs::dlg_message(paste0("there are ", length(updatedFileList), " mzml files in ", length(plateList), " plates"))


#get list of dirs
dirPath <- list.dirs(filePath)
dirPath <- dirPath[!grepl("(?i)archive", dirPath)]
dirPath <- dirPath[-which(dirPath %in% filePath)]

#organise into sub-plates (a and b)
idxDir = dirPath[1] #for loop testing
for(idxDir in dirPath){
 
   #only run next code if parent plates remain (i.e. it hasn't been sorted into a and b sub plates yet
  if(gsub(".*/", "", idxDir) %in% plateList){

#get list of filenames
fileName <-
  tibble(
    oldName = list.files(path = idxDir,
                         pattern = "(.?)mzML")
  )

#append injection number
fileName <- fileName %>%
  add_column(
    numbers = gsub(".*_", "", fileName$oldName) %>% gsub("\\..*", "", .)
  )

#remove rerun tag
fileName$numbers <- gsub("-.*", "", fileName$numbers)

#add 00
for(idx in 1:length(fileName$numbers)){
  if(nchar(fileName$numbers[idx]) ==1){fileName$numbers[idx] <- paste0("00", fileName$numbers[idx])}
  if(nchar(fileName$numbers[idx]) ==2){fileName$numbers[idx] <- paste0("0", fileName$numbers[idx])}
}

fileName[["newName"]] <- paste0(fileName$numbers, "_", fileName$oldName)

#run for loop to rename
for(idxName in fileName$oldName){
  file.rename(
    from=paste0(idxDir, "/", idxName),
    to=paste0(idxDir, "/", fileName$newName[which(fileName$oldName==idxName)])
  )
}

#sort file name by newName
fileName <- fileName %>% arrange(newName)

#identify the check solutions
checkFileName <- fileName %>%
  filter(grepl("(?i)check", oldName)) %>%
  arrange(newName)

#find check index
checkIdx <- which(fileName$newName %in% checkFileName$newName)

  #set up plate a
  dir.create(paste0(idxDir, "a"))

  #move files from parent folder to folder a
  for(idxName in fileName$newName[1:checkIdx[2]]){
    file.rename(
      from=paste0(idxDir, "/", idxName),
      to=paste0(idxDir, "a/", idxName)
    )
  }

  if(length(checkIdx) == 3){
    #set up plate b
    dir.create(paste0(idxDir, "b"))

    #move files from parent folder to folder b
    for(idxName in fileName$newName[(checkIdx[2]+1):nrow(fileName)]){
      file.rename(
        from=paste0(idxDir, "/", idxName),
        to=paste0(idxDir, "b/", idxName)
      )
    }

    #copy the mid check file so both a and b have a copy
    file.copy(
      from=paste0(idxDir, "a/", checkFileName$newName[2]),
      to=paste0(idxDir, "b/", checkFileName$newName[2])
    )
  }

  #tidy up and remove original folder
  #unlink(idxDir, recursive=TRUE)

  }
}
}

svDialogs::dlg_message("check folders. If all ahs worked correctly delete the original folders (without the a or b sub-tag)")



```

## peform directory check

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

updatedFileList <- list.files(filePath, pattern = ".mzML", recursive = TRUE, full.names = TRUE); updatedFileList <- updatedFileList[!grepl("archive",updatedFileList)]; length(updatedFileList)

#get list of dirs
dirPath <- list.dirs(filePath)
dirPath <- dirPath[!grepl("(?i)archive", dirPath)]
dirPath <- dirPath[-which(dirPath %in% filePath)]

svDialogs::dlg_message(paste0("there are ", length(updatedFileList), " mzml files in ", length(dirPath), " sub-plates (a/b): ", paste0(gsub(".*/", "", dirPath), collapse = "; ")))

svDialogs::dlg_message(paste0("if this is incorrect, finalise subfoders and rerun this chunk"))

```


## read in templates

## prepare mrm method list from a msFile text export

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}


#method copy
svDialogs::dlg_message(paste0("copy the AB and N .exp methods used into a project methods folder: ", gsub(".*/", "~/", master_list$project_details$project_dir), "/msMethods"))


listMethods <- list.files(path = paste0(master_list$project_details$project_dir, '/msMethods/'), pattern = ".exp", full.names = TRUE)
methodPath <- listMethods[grepl(paste0("_",master_list$project_details$wash,"_"), listMethods)]

#read in delim ms method data from a waters .exp file
msDataMethod <- read.table(
  file = methodPath,
  header = FALSE, 
  sep = " ", 
  col.names = paste0("V",seq_len(20)), 
  fill = TRUE) %>% 
  as_tibble() %>%
  add_column(
    idx = c(1:nrow(.)),
    .before = 1
  )

#extract function list
functionList <- bind_cols(
  msDataMethod %>% filter(grepl("FUNCTION", V1)) %>% unite(mrmFunction, V1, V2) %>% select(idx, mrmFunction),
  tibble(
    method = msDataMethod %>% filter(grepl("(?i)CompoundName_1", V1)) %>% .$V1 %>% gsub(".*\\,", "", .),
    experiment = msDataMethod %>% filter(grepl("(?i)CompoundName_1", V1)) %>% .$V2, #%>% gsub(".*,", " ", .),
    metabolite = msDataMethod %>% filter(grepl("(?i)CompoundName_1", V1)) %>% .$V3 %>% gsub("\\..*", "", .),
    #metabolite = msDataMethod %>% filter(grepl("(?i)CompoundName_1", V1)) %>%  
    #   mutate(V2 = gsub(",", "-", V2)) %>%
    #   mutate(V1 = gsub(".*,", "_ _,", V1)) %>%
    #   paste0("_1_", .$V1, "_2_", .$V2, "_3_") %>%
    #   sub(".*,", "", .) %>%
    #   sub("_2__3_", "", .),
    rtStart = msDataMethod %>% filter(grepl("(?i)FunctionStartTime", V1)) %>% .$V1 %>% gsub(".*,", "", .),
    rtEnd = msDataMethod %>% filter(grepl("(?i)FunctionEndTime", V1)) %>% .$V1 %>% gsub(".*,", "", .),
    polarity = msDataMethod %>% filter(grepl("(?i)FunctionPolarity", V1)) %>% .$V1 %>% gsub(".*,", "", .),
    mrm1_q1 = msDataMethod %>% filter(grepl("(?i)SIRMass1", V1)) %>% .$V1 %>% gsub(".*,", "", .) %>% as.numeric(),
    mrm1_q3 = msDataMethod %>% filter(grepl("(?i)SIRMass_2_1", V1)) %>% .$V1 %>% gsub(".*,", "", .) %>% as.numeric(),
    mrm1_ce = msDataMethod %>% filter(grepl("(?i)collisionenergy\\(V\\)_1", V1)) %>% .$V1 %>% gsub(".*,", "", .) %>% as.numeric(),
    mrm2_q1 = NA,
    mrm2_q3 = NA,
    mrm2_ce = NA,
    mrm3_q1 = NA,
    mrm3_q3 = NA,
    mrm3_ce = NA
  )
) %>% 
  # create experiment channel
  #add_column(
   # .before = "metabolite",
  #  experiment = sub("_2_.*", "", .$metabolite)
  #) %>%
  mutate(metabolite = sub(".*_2_", "", metabolite)) %>%
  mutate(metabolite = sub("_3_.*", "", metabolite)) %>%
  mutate(experiment = sub("_3_.*", "", experiment)) %>%
  mutate(experiment = sub(".*-QC", "QC", experiment)) %>%
  mutate(metabolite = sub("-QC", "", metabolite)) %>%
  mutate(experiment = sub("_EXP", "", experiment)) %>%
  mutate(experiment = sub("\\(.*", "", experiment)) %>%
  mutate(experiment = sub("Pphen", "PPHEN", experiment)) %>%
  mutate(experiment = sub("PPh", "PPHEN", experiment)) %>%
  mutate(experiment = sub("Pphen", "PPHEN", experiment))
  

#create indole experiment
functionList$experiment[which(functionList$metabolite == "IBA")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "ICA")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "ICARB")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "I3AA")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "I3AA-SIL_d4")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "IAAEE")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "IAAME")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "IACA")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "IACR")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "ILA")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "IET")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "5-HIAA")] <- "INDOLE"
functionList$experiment[which(functionList$metabolite == "5-HIAA-SIL_d5")] <- "INDOLE"


#tidy up method file by adding blank space as experiment where extraction of info has an error
for(idx_metabolite in functionList$metabolite){
  functionList[["experiment"]][which(functionList[["experiment"]]==idx_metabolite)] <- ""
}

#add experiment type for each SIL
for(idx_sil in functionList$metabolite[grepl("-SIL", functionList$metabolite)]){
  experiment <- functionList[["experiment"]][which(functionList[["metabolite"]]==gsub("-SIL.*","",idx_sil))]
  if(length(experiment) ==1){
    functionList[["experiment"]][which(functionList[["metabolite"]]==idx_sil)] <- experiment
  }
}


#add Q2 and Q3 (missmatch in length so can't do simple string extract)
idxStart = functionList$idx[1]
for(idxStart in functionList$idx){
  
  #find end index of function data
  idxEnd <- functionList$idx[which(functionList$idx == idxStart)+1]
  
  #extract data specific for function
  msDataLoop <- msDataMethod %>%
    filter(idx >= idxStart & idx <= idxEnd)
  
  #extract mrm2/mrm3 data
  mrm2_q1 <- msDataLoop %>% filter(grepl("(?i)SIRMass2", V1)) %>% .$V1 %>% gsub(".*,", "", .) %>% as.numeric()
  mrm2_q3 <- msDataLoop %>% filter(grepl("(?i)SIRMass_2_2", V1)) %>% .$V1 %>% gsub(".*,", "", .) %>% as.numeric()
  mrm2_ce <- msDataLoop %>% filter(grepl("(?i)collisionenergy\\(V\\)_2", V1)) %>% .$V1 %>% gsub(".*,", "", .) %>% as.numeric()
  mrm3_q1 <- msDataLoop %>% filter(grepl("(?i)SIRMass3", V1)) %>% .$V1 %>% gsub(".*,", "", .) %>% as.numeric()
  mrm3_q3 <- msDataLoop %>% filter(grepl("(?i)SIRMass_2_3", V1)) %>% .$V1 %>% gsub(".*,", "", .) %>% as.numeric()
  mrm3_ce <- msDataLoop %>% filter(grepl("(?i)collisionenergy\\(V\\)_3", V1)) %>% .$V1 %>% gsub(".*,", "", .) %>% as.numeric()
  
  #fill columns
  if(length(mrm2_q1)==1){functionList$mrm2_q1[which(functionList$idx == idxStart)] <- mrm2_q1}
  if(length(mrm2_q3)==1){functionList$mrm2_q3[which(functionList$idx == idxStart)] <- mrm2_q3}
  if(length(mrm2_ce)==1){functionList$mrm2_ce[which(functionList$idx == idxStart)] <- mrm2_ce}
  if(length(mrm3_q1)==1){functionList$mrm3_q1[which(functionList$idx == idxStart)] <- mrm3_q1}
  if(length(mrm3_q3)==1){functionList$mrm3_q3[which(functionList$idx == idxStart)] <- mrm3_q3}
  if(length(mrm3_ce)==1){functionList$mrm3_ce[which(functionList$idx == idxStart)] <- mrm3_ce}
}


#trim method if running control charts to only matched SILs

functionListMethod <- "CCC/fullAnalysis"

while(functionListMethod != "CCC" & functionListMethod != "fullAnalysis"){
functionListMethod <- svDialogs::dlg_input(message = paste0("are you running a controlChartCheck or a full project analysis?"), default = paste0("CCC/fullAnalysis"))$res
}

if(functionListMethod == "CCC"){
  
  methodSil <- functionList$metabolite[grepl("-SIL", functionList$metabolite)]
  matchedSil <- methodSil %>% gsub("-SIL.*", "", .)
  matchedSil <- functionList$metabolite[functionList$metabolite %in% matchedSil]
  
  nList <- c(
     "5-HIAA", "5-HIAA-SIL_d5", "CDCA", "CDCA-SIL_d4", "GCDCA", "GCDCA-SIL_d4", "GUDCA", "GUDCA-SIL_d4", 
    "XA", "XA-SIL_d4", "12\\(R\\)-HETE", "12\\(R\\)-HETE-SIL_d8",  "15\\(S\\)-HETE","15\\(S\\)-HETE-SIL_d8", "8i-PGF2a", "8i-PGF2a-SIL_d4"
    )
  
  abList <- c(
    "3-HAA", "3-HAA-SIL_D3", "CAR\\(2:0\\)", "CAR\\(2:0\\)-SIL_d3", "CAR\\(4\\:0\\)", "CAR\\(4\\:0\\)-SIL_d3", "CAR\\(i5\\:0\\)", "CAR\\(i5\\:0\\)-SIL_d9", "KYN", "KYN-SIL_D4", "NEO", "NEO-SIL_13C5", "TRYP", "TRYP-SIL_d5", "SER", "SER-SIL_d4"
  )
  
functionList <- functionList %>%
  filter(metabolite %in% c(methodSil, matchedSil))

  if(master_list$project_details$wash[1] == "N"){
    functionList <- functionList %>%
    filter(grepl(paste0("^(", paste(nList, collapse = "|"), ")$"), metabolite))
  }

 if(master_list$project_details$wash[1] == "AB"){
    functionList <- functionList %>%
    filter(grepl(paste0("^(", paste(abList, collapse = "|"), ")$"), metabolite))
  }

}



```

## review mrm method

This section offers the user the opportunity to review the MS method to make sure the conversion from the .exp file has worked appropiately. Please review the method and make any changes to the metabolite and experiment columns.

```{r}

#export method
write_csv(functionList,
          paste0(master_list$project_details$project_dir,
                 "/msMethods/",
                 Sys.Date(),
                 "_", master_list$project_details$wash,
                 "_method_forReview.csv"
          ))


svDialogs::dlg_message("now review method for errors (save in method subfolder)")

svDialogs::dlg_message(paste0("review method, make any ad-hoc edits in the experiment and metabolite columns, then rename to _method_reviewed_", master_list$project_details$user_name, ".csv"))


methodUpdateFile = list.files(
  path = paste0(master_list$project_details$project_dir,"/msMethods/"), 
  pattern = paste0("_", master_list$project_details$wash, "_method_reviewed_")
  )

svDialogs::dlg_message(paste0("is this your reviewed file ....................................................................... ", 
                              methodUpdateFile, " ..... ...... ............ ................... .......... ........... .......... ............ .......... ..................... ......... .......... if not rerun chunk and edit name of file. move old files to archive folder."))



#store in masterlist
master_list$templates[['msMethod']] <- read_csv(
  paste0(master_list$project_details$project_dir,
                 "/msMethods/",
                 methodUpdateFile
  )
)

svDialogs::dlg_message("if you notice any method errors in experiments/metabolite names columns - update method version on the instrument before next project")

```

## mrm method pivot_longer

```{r}

#mke long table
master_list$templates[['msMethodLong']] <- left_join(
  by="rowIdx",
  master_list$templates[['msMethod']] %>%
    select(idx, mrmFunction, metabolite, rtStart, rtEnd, polarity, mrm1_q1, mrm2_q1, mrm3_q1) %>%
    pivot_longer(cols = c(mrm1_q1, mrm2_q1, mrm3_q1), 
                 names_to = "mrm",
                 values_to = "q1") %>%
    add_column(rowIdx = (1:nrow(.)), .before = 1),
  master_list$templates[['msMethod']] %>%
    select(idx, mrm1_q3, mrm2_q3, mrm3_q3) %>%
    pivot_longer(cols = c(mrm1_q3, mrm2_q3, mrm3_q3), 
                 names_to = "mrm",
                 values_to = "q3") %>%
    add_column(rowIdx = (1:nrow(.)), .before = 1) %>%
    select(-mrm, -idx),
)
  
#tidy up mrm column
master_list$templates[['msMethodLong']]$mrm <- gsub("_.*", "", master_list$templates[['msMethodLong']]$mrm)

#remove NA mrms
master_list$templates[['msMethodLong']] <- master_list$templates[['msMethodLong']][-intersect(which(is.na(master_list$templates[['msMethodLong']]$q1)), which(is.na(master_list$templates[['msMethodLong']]$q3))),]


```

## create list of plates for analysis

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

############
#create plate list 
#set filePath
filePath <- paste(sep="/", master_list$project_details$project_dir,  master_list$project_details$wash, "data/mzml")

#get list of dirs
dirPath <- list.dirs(filePath)
dirPath <- dirPath[!grepl("(?i)archive", dirPath)]
dirPath <- dirPath[-which(dirPath %in% filePath)]

#store in masterList
master_list$project_details$plateDirList <- dirPath

#get plate list
master_list$project_details$plateList <- gsub(".*\\/", "", master_list$project_details$plateDirList)

svDialogs::dlg_message(paste0("plates for processing: ", paste0(collapse = "; ", master_list$project_details$plateList)))

```

## list mzml

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

# PROCESS: create list of mzML FILES USING mzR ------------------------------------
#idx_plate = master_list$project_details$plateList[1] #for testing loop
for(idx_plate in master_list$project_details$plateList){
  
  #create empty list for storing mzR objects
  master_list$data$mzR[[idx_plate]] <- list()
  
  #get list of mzMl files
  master_list$data$mzR[[idx_plate]] <- list.files(
    pattern = "(?i).mzml",
    paste0(
      master_list$project_details$plateDirList[grepl(idx_plate, master_list$project_details$plateDirList)]
      )    
    )

  #remove conditioning runs
master_list$data$mzR[[idx_plate]] <- master_list$data$mzR[[idx_plate]][!grepl("(?i)cond", master_list$data$mzR[[idx_plate]])]
}
  
```

## Find peak boundaries for targets

identify peak boundaries using check solution as a reference. 

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

master_list$data$peakRegion <- list()
#set object to temp store checkMzMLs
checkMzml <- list()

idx_plate = master_list$project_details$plateList[1] #for testing
#for(idx_plate in master_list$project_details$plateList){
for(idx_plate in master_list$project_details$plateList){
  
  print(idx_plate)
  
  #find index of check solution within idx_plate mzml directory
  checkIndex <- master_list$data$mzR[[idx_plate]][grep("check", master_list$data$mzR[[idx_plate]], ignore.case = TRUE)]
  
  idx_mzML = checkIndex[1] #for testing loop
  for(idx_mzML in checkIndex){
    print(idx_mzML)
    
    #read in mzML
    if(!gsub("(?i).mzml", "", idx_mzML) %in% names(checkMzml)){
      #create mzR object
      checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_object <- mzR::openMSfile(
        filename = paste0(master_list$project_details$plateDirList[grepl(idx_plate, master_list$project_details$plateDirList)], "/", idx_mzML)
      )
      #create mzR header
      checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_header <- mzR::chromatogramHeader(
        checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_object) %>%
        as_tibble() %>% 
        #mutate to avoid low value decimal making error in the output
        mutate(precursorIsolationWindowTargetMZ = round(precursorIsolationWindowTargetMZ, 4),
               productIsolationWindowTargetMZ = round(productIsolationWindowTargetMZ, 4))
      #create mzR chromatogram
      checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_chromatogram <- mzR::chromatograms(
        checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_object)
    }
    
    
    #cycle through each function to extract the chrom data
    #for testing
    idx_function = "FUNCTION_13"
    
    for(idx_function in unique(master_list$templates$msMethodLong %>% filter(!grepl("-SIL", metabolite)) %>% .$mrmFunction)){
      
       #get metabolite string
       mrm_metabolite = master_list$templates$msMethodLong %>%
           filter(mrmFunction == idx_function) %>%
           .$metabolite %>%
           unique()
       
       
      #get chromatogram header data
      mzmlFuncHeader <- checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_header %>%
        filter(grepl(paste0(gsub("_","=", casefold(idx_function)), " "), chromatogramId))
      
      
      #do a check in case the functions have been updated in the project and change between files
      
      if(FALSE %in% (c(
          mzmlFuncHeader$precursorIsolationWindowTargetMZ %in% (master_list$templates$msMethodLong %>% filter(mrmFunction == idx_function) %>% .$q1 %>% unique()),
          mzmlFuncHeader$productIsolationWindowTargetMZ %in% (master_list$templates$msMethodLong %>% filter(mrmFunction == idx_function) %>% .$q3 %>% unique())
        ))){
        
        #update idx_function
        idx_function = checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_header %>% 
          filter(precursorIsolationWindowTargetMZ %in% (master_list$templates$msMethodLong %>% filter(mrmFunction == idx_function) %>% .$q1 %>% unique())) %>% 
          filter(productIsolationWindowTargetMZ %in% (master_list$templates$msMethodLong %>% filter(mrmFunction == idx_function) %>% .$q3 %>% unique())) %>%
          .$chromatogramId %>%
          gsub(".*function=", "function=", . ) %>%
          gsub(" .*", "", .) %>%
          gsub("=", "_", .) %>% 
          toupper() %>%
          unique()
        
        #update header
        mzmlFuncHeader <- checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_header %>%
          filter(grepl(paste0(gsub("_","=", casefold(idx_function)), " "), chromatogramId))
        
        }
      
      
     #set channel index (within msFunction) for storing later
     channel_index = 1
       
      idx_mrm = mzmlFuncHeader$chromatogramIndex[1]#for testing
      for(idx_mrm in mzmlFuncHeader$chromatogramIndex){
      #get metabolite metadata   
      row_index = which(mzmlFuncHeader$chromatogramIndex==idx_mrm)
      #store mrm details
      precursor_mz = mzmlFuncHeader$precursorIsolationWindowTargetMZ[row_index]
      product_mz = mzmlFuncHeader$productIsolationWindowTargetMZ[row_index]
      
        #set empty object for storing peak boundaries
      if(!mrm_metabolite %in% names(master_list$data$peakRegion)){
        master_list$data$peakRegion[[mrm_metabolite]] <- list()
      } 
      
      #set modifiable loop chromatogram
      loopChromatogram <- checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_chromatogram[[idx_mrm]] %>%
        as_tibble() %>%
        add_column(.before = 1,
          scanIndex = 1:nrow(.)
        ) %>%
        setNames(c("scanIndex", "time", "intensity"))
      
      #create peak table
      peakTable = tibble()
      #fl.count = 9
      smoothData <- loopChromatogram %>%
          select(scanIndex, time) %>%
          add_column(
          intensity = loopChromatogram$intensity %>%
            pracma::savgol(fl = 7, forder = 0, dorder = 0) %>%
            pracma::savgol(fl = 5, forder = 0, dorder = 0) %>%
            pracma::savgol(fl = 3, forder = 0, dorder = 0)
            ) 
      
      smoothData <- smoothData %>%
        as.matrix() %>%
        smoothr::smooth_chaikin(., wrap = FALSE) %>%
        as_tibble() %>%
        setNames(c("scanIndex", "time", "intensity"))

        #reset baseline to help findPeaks
        smoothData$intensity[which(smoothData$intensity < 
                                     (sd(smoothData$intensity)*0.1))] <- min(smoothData$intensity)
          #create peak table
      peakTable <- smoothData[["intensity"]] %>%
        pracma::findpeaks(x=., nups = 5, ndowns = 5,  minpeakdistance = 20, sortstr = TRUE, minpeakheight = sd(.)) %>%
        as_tibble() %>% setNames(c("intensity", "peakApexIndex", "peakStartIndex", "peakEndIndex"))
      
      if(nrow(peakTable) == 0){
        peakTable <- tibble(
          intensity = max(smoothData$intensity),
          peakApexIndex = which.max(smoothData$intensity),
          peakStartIndex = which(smoothData$intensity > min(smoothData$intensity))[1],
          peakEndIndex = rev(which(smoothData$intensity > min(smoothData$intensity)))[1]
          )
      }

      #set index of which peak to use from peakDetect table
      peakTableRow = 1
      
      #store peak details
      peakApexInt <- peakTable[["intensity"]][peakTableRow]
      peakApexIndex <- peakTable[["peakApexIndex"]][peakTableRow]; peakApexRT <- smoothData[["time"]][peakApexIndex]
      peakStartIndex <- peakTable[["peakStartIndex"]][peakTableRow]; peakStartRT <- smoothData[["time"]][peakStartIndex]
      peakEndIndex <- peakTable[["peakEndIndex"]][peakTableRow]; peakEndRT <- smoothData[["time"]][peakEndIndex]
      peakWidth <- peakEndIndex-peakStartIndex
      
       #creat a tibble to define peak
      master_list$data$peakRegion[[mrm_metabolite]] <- bind_rows(
        master_list$data$peakRegion[[mrm_metabolite]],
        tibble(
          ms_function = idx_function,
          ms_channel = channel_index,
          ms_header_id = idx_mrm,
          sampleMetabolite = mrm_metabolite,
          mrm = paste0(round(precursor_mz,2), " > ", round(product_mz,2)),
          precursor_mz = precursor_mz,
          product_mz = product_mz,
          samplePlate = idx_plate,
          sampleName = idx_mzML,
          peakApexInt = peakApexInt,
          peakStartRT = peakStartRT,
          peakApexRT = peakApexRT,
          peakEndRT = peakEndRT,
          peakStartIndex =peakStartIndex,
          peakApexIndex = peakApexIndex,
          peakEndIndex =peakEndIndex,
          peakWidth = peakWidth,
          #flCount = fl.count
        )
      )
      
      channel_index <- channel_index+1
      
       } #idx_mrm
      
    } #idx_function
    
  } #idx_mzML
  
} #idx_plate
    
    


## finalise peak boundaries

#identify Q and q ions and finalise boundaries per plate

master_list$data$peakRegionFinal <- list()

#idx_metabolite = names(master_list$data$peakRegion)[13]#for testing loop
#idx_metabolite = "LTE4"
for(idx_metabolite in names(master_list$data$peakRegion)){
  print(idx_metabolite)
  
  master_list$data$peakRegionFinal[[idx_metabolite]] <- list()
  
  #first ID Q ion across plate
    q_tibble <- list()
    
    idx_channel = unique(master_list$data$peakRegion[[idx_metabolite]]$ms_channel)[1]
    for(idx_channel in unique(master_list$data$peakRegion[[idx_metabolite]]$ms_channel)){
      mean_intensity = master_list$data$peakRegion[[idx_metabolite]] %>%
        filter(ms_channel == idx_channel) %>%
        .$peakApexInt %>%
        median(na.rm = TRUE)
      q_tibble <- bind_rows(
        q_tibble,
        tibble(
          ms_channel = idx_channel,
          mean_int = mean_intensity)
      )
    }
  
  #identify Q ion channel 
  Q_ion <- q_tibble$ms_channel[which.max(q_tibble$mean_int)]
  
  master_list$data$peakRegion[[idx_metabolite]]$Q_q <- "q"
  master_list$data$peakRegion[[idx_metabolite]]$Q_q[which(master_list$data$peakRegion[[idx_metabolite]]$ms_channel == Q_ion)] <- "Q"
  
  
  #finalise peak bounday of Q ion
  idx_plate = unique(master_list$data$peakRegion[[idx_metabolite]]$samplePlate)[1] #for testing loop
  #for(idx_plate in unique(master_list$data$peakRegion[[idx_metabolite]]$samplePlate)[1]){ #for testing loop
  for(idx_plate in unique(master_list$data$peakRegion[[idx_metabolite]]$samplePlate)){
    print(idx_plate)
    plate_data <- master_list$data$peakRegion[[idx_metabolite]] %>%
      filter(samplePlate==idx_plate) %>%
      filter(Q_q == "Q")
    
    master_list$data$peakRegionFinal[[idx_metabolite]] <- bind_rows(
      master_list$data$peakRegionFinal[[idx_metabolite]],
      tibble(
        sample_plate = idx_plate,
        ms_function = plate_data$ms_function %>% unique(),
        ms_channel_Q = plate_data$ms_channel %>% unique(),
        ms_header_id = plate_data$ms_header_id %>% unique(),
        metabolite = idx_metabolite,
        precursor_mz = plate_data$precursor_mz %>% unique() %>% round(2),
        product_mz = plate_data$product_mz %>% unique() %>% round(2),
        mrm = paste0(plate_data$precursor_mz %>% unique() %>% round(2), " > " , plate_data$product_mz %>% unique() %>% round(2)),
        peakStartRT = (((plate_data$peakStartRT %>% min(na.rm = TRUE)) *100) %>% floor()) /100,
        peakEndRT =  (((plate_data$peakEndRT %>% min(na.rm = TRUE)) *100) %>% ceiling()) /100,
        peakApexRT = plate_data$peakApexRT %>% mean(na.rm = TRUE) %>% round(2),
        peakStartIndex = plate_data$peakStartIndex %>% min(na.rm = TRUE) %>% round(2),
        peakEndIndex = plate_data$peakEndIndex %>% max(na.rm = TRUE) %>% round(2)
      )
    )
  
  }#idx_plate
}#idx_metabolite


```


## Find peak boundaries for SIL IS

identify peak boundaries using LTR as a reference as internal standards not in check solution. 

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

master_list$data$peakRegionSIL <- list()
#set object to temp store checkMzMLs
checkMzml <- list()

idx_plate = master_list$project_details$plateList[1] #for testing
#for(idx_plate in master_list$project_details$plateList){
for(idx_plate in master_list$project_details$plateList){
  
  #find index of check solution within idx_plate mzml directory
  checkIndex <- master_list$data$mzR[[idx_plate]][grep("vLTR", master_list$data$mzR[[idx_plate]], ignore.case = TRUE)]
  
  idx_mzML = checkIndex[1] #for testing loop
  for(idx_mzML in checkIndex){
    
    #read in mzML
    if(!gsub("(?i).mzml", "", idx_mzML) %in% names(checkMzml)){
      #create mzR object
      checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_object <- mzR::openMSfile(
        filename = paste0(master_list$project_details$plateDirList[grepl(idx_plate, master_list$project_details$plateDirList)], "/", idx_mzML)
      )
      #create mzR header
      checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_header <- mzR::chromatogramHeader(
        checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_object) %>%
        as_tibble() %>% 
        #mutate to avoid low value decimal making error in the output
        mutate(precursorIsolationWindowTargetMZ = round(precursorIsolationWindowTargetMZ, 4),
               productIsolationWindowTargetMZ = round(productIsolationWindowTargetMZ, 4))
      #create mzR chromatogram
      checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_chromatogram <- mzR::chromatograms(
        checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_object)
    }
    
    
    #cycle through each function to extract the chrom data
    #for testing
    idx_function = "FUNCTION_15"
    
    for(idx_function in unique(master_list$templates$msMethodLong %>% filter(grepl("SIL", metabolite)) %>% .$mrmFunction)){
      
       #get metabolite string
       mrm_metabolite = master_list$templates$msMethodLong %>%
           filter(mrmFunction == idx_function) %>%
           .$metabolite %>%
           unique()
       
       
      #get chromatogram header data
      mzmlFuncHeader <- checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_header %>%
        filter(grepl(paste0(gsub("_","=", casefold(idx_function)), " "), chromatogramId))
      
      
      #do a check in case the functions have been updated in the project and change between files
      
      if(FALSE %in% (c(
          mzmlFuncHeader$precursorIsolationWindowTargetMZ %in% (master_list$templates$msMethodLong %>% filter(mrmFunction == idx_function) %>% .$q1 %>% unique()),
          mzmlFuncHeader$productIsolationWindowTargetMZ %in% (master_list$templates$msMethodLong %>% filter(mrmFunction == idx_function) %>% .$q3 %>% unique())
        ))){
        
        #update idx_function
        idx_function = checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_header %>% 
          filter(precursorIsolationWindowTargetMZ %in% (master_list$templates$msMethodLong %>% filter(mrmFunction == idx_function) %>% .$q1 %>% unique())) %>% 
          filter(productIsolationWindowTargetMZ %in% (master_list$templates$msMethodLong %>% filter(mrmFunction == idx_function) %>% .$q3 %>% unique())) %>%
          .$chromatogramId %>%
          gsub(".*function=", "function=", . ) %>%
          gsub(" .*", "", .) %>%
          gsub("=", "_", .) %>% 
          toupper() %>%
          unique()
        
        #update header
        mzmlFuncHeader <- checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_header %>%
          filter(grepl(paste0(gsub("_","=", casefold(idx_function)), " "), chromatogramId))
        
        }
      
      
     #set channel index (within msFunction) for storing later
     channel_index = 1
       
      idx_mrm = mzmlFuncHeader$chromatogramIndex[1]#for testing
      for(idx_mrm in mzmlFuncHeader$chromatogramIndex){
      #get metabolite metadata   
      row_index = which(mzmlFuncHeader$chromatogramIndex==idx_mrm)
      #store mrm details
      precursor_mz = mzmlFuncHeader$precursorIsolationWindowTargetMZ[row_index]
      product_mz = mzmlFuncHeader$productIsolationWindowTargetMZ[row_index]
      
        #set empty object for storing peak boundaries
      if(!mrm_metabolite %in% names(master_list$data$peakRegionSIL)){
        master_list$data$peakRegionSIL[[mrm_metabolite]] <- list()
      } 
      
      #set modifiable loop chromatogram
      loopChromatogram <- checkMzml[[gsub("(?i).mzml", "", idx_mzML)]]$mzR_chromatogram[[idx_mrm]] %>%
        as_tibble() %>%
        add_column(.before = 1,
          scanIndex = 1:nrow(.)
        ) %>%
        setNames(c("scanIndex", "time", "intensity"))
      
      #create peak table
      peakTable = tibble()
      #fl.count = 9
      smoothData <- loopChromatogram %>%
          select(scanIndex, time) %>%
          add_column(
          intensity = loopChromatogram$intensity %>%
            pracma::savgol(fl = 7, forder = 0, dorder = 0) %>%
            pracma::savgol(fl = 5, forder = 0, dorder = 0) %>%
            pracma::savgol(fl = 3, forder = 0, dorder = 0)
            ) 
      
      smoothData <- smoothData %>%
        as.matrix() %>%
        smoothr::smooth_chaikin(., wrap = FALSE) %>%
        as_tibble() %>%
        setNames(c("scanIndex", "time", "intensity"))

        #reset baseline to help findPeaks
        smoothData$intensity[which(smoothData$intensity < 
                                     (sd(smoothData$intensity)*0.1))] <- min(smoothData$intensity)
          #create peak table
      peakTable <- smoothData[["intensity"]] %>%
        pracma::findpeaks(x=., nups = 5, ndowns = 5,  minpeakdistance = 20, sortstr = TRUE, minpeakheight = sd(.)) %>%
        as_tibble() %>% setNames(c("intensity", "peakApexIndex", "peakStartIndex", "peakEndIndex"))
      
      if(nrow(peakTable) == 0){
        peakTable <- tibble(
          intensity = max(smoothData$intensity),
          peakApexIndex = which.max(smoothData$intensity),
          peakStartIndex = which(smoothData$intensity > min(smoothData$intensity))[1],
          peakEndIndex = rev(which(smoothData$intensity > min(smoothData$intensity)))[1]
          )
      }

      #set index of which peak to use from peakDetect table
      peakTableRow = 1
      
      #store peak details
      peakApexInt <- peakTable[["intensity"]][peakTableRow]
      peakApexIndex <- peakTable[["peakApexIndex"]][peakTableRow]; peakApexRT <- smoothData[["time"]][peakApexIndex]
      peakStartIndex <- peakTable[["peakStartIndex"]][peakTableRow]; peakStartRT <- smoothData[["time"]][peakStartIndex]
      peakEndIndex <- peakTable[["peakEndIndex"]][peakTableRow]; peakEndRT <- smoothData[["time"]][peakEndIndex]
      peakWidth <- peakEndIndex-peakStartIndex
      
       #creat a tibble to define peak
      master_list$data$peakRegionSIL[[mrm_metabolite]] <- bind_rows(
        master_list$data$peakRegionSIL[[mrm_metabolite]],
        tibble(
          ms_function = idx_function,
          ms_channel = channel_index,
          ms_header_id = idx_mrm,
          sampleMetabolite = mrm_metabolite,
          mrm = paste0(round(precursor_mz,2), " > ", round(product_mz,2)),
          precursor_mz = precursor_mz,
          product_mz = product_mz,
          samplePlate = idx_plate,
          sampleName = idx_mzML,
          peakApexInt = peakApexInt,
          peakStartRT = peakStartRT,
          peakApexRT = peakApexRT,
          peakEndRT = peakEndRT,
          peakStartIndex =peakStartIndex,
          peakApexIndex = peakApexIndex,
          peakEndIndex =peakEndIndex,
          peakWidth = peakWidth,
          #flCount = fl.count
        )
      )
      
      channel_index <- channel_index+1
      
       } #idx_mrm
      
    } #idx_function
    
  } #idx_mzML
  
} #idx_plate
    
    


## finalise peak boundaries

#identify Q and q ions and finalise boundaries per plate

master_list$data$peakRegionSILFinal <- list()

#idx_metabolite = names(master_list$data$peakRegionSIL)[13]#for testing loop
#idx_metabolite = "LTE4"
for(idx_metabolite in names(master_list$data$peakRegionSIL)){
  
  master_list$data$peakRegionSILFinal[[idx_metabolite]] <- list()
  
  #first ID Q ion across plate
    q_tibble <- list()
    
    idx_channel = unique(master_list$data$peakRegionSIL[[idx_metabolite]]$ms_channel)[1]
    for(idx_channel in unique(master_list$data$peakRegionSIL[[idx_metabolite]]$ms_channel)){
      mean_intensity = master_list$data$peakRegionSIL[[idx_metabolite]] %>%
        filter(ms_channel == idx_channel) %>%
        .$peakApexInt %>%
        median(na.rm = TRUE)
      q_tibble <- bind_rows(
        q_tibble,
        tibble(
          ms_channel = idx_channel,
          mean_int = mean_intensity)
      )
    }
  
  #identify Q ion channel 
  Q_ion <- q_tibble$ms_channel[which.max(q_tibble$mean_int)]
  
  master_list$data$peakRegionSIL[[idx_metabolite]]$Q_q <- "q"
  master_list$data$peakRegionSIL[[idx_metabolite]]$Q_q[which(master_list$data$peakRegionSIL[[idx_metabolite]]$ms_channel == Q_ion)] <- "Q"
  
  
  #finalise peak bounday of Q ion
  idx_plate = unique(master_list$data$peakRegionSIL[[idx_metabolite]]$samplePlate)[1] #for testing loop
  #for(idx_plate in unique(master_list$data$peakRegionSIL[[idx_metabolite]]$samplePlate)[1]){ #for testing loop
  for(idx_plate in unique(master_list$data$peakRegionSIL[[idx_metabolite]]$samplePlate)){
    plate_data <- master_list$data$peakRegionSIL[[idx_metabolite]] %>%
      filter(samplePlate==idx_plate) %>%
      filter(Q_q == "Q")
    
    master_list$data$peakRegionSILFinal[[idx_metabolite]] <- bind_rows(
      master_list$data$peakRegionSILFinal[[idx_metabolite]],
      tibble(
        sample_plate = idx_plate,
        ms_function = plate_data$ms_function %>% unique(),
        ms_channel_Q = plate_data$ms_channel %>% unique(),
        ms_header_id = plate_data$ms_header_id %>% unique(),
        metabolite = idx_metabolite,
        precursor_mz = plate_data$precursor_mz %>% unique() %>% round(2),
        product_mz = plate_data$product_mz %>% unique() %>% round(2),
        mrm = paste0(plate_data$precursor_mz %>% unique() %>% round(2), " > " , plate_data$product_mz %>% unique() %>% round(2)),
        peakStartRT = (((plate_data$peakStartRT %>% min(na.rm = TRUE)) *100) %>% floor()) /100,
        peakEndRT =  (((plate_data$peakEndRT %>% min(na.rm = TRUE)) *100) %>% ceiling()) /100,
        peakApexRT = plate_data$peakApexRT %>% mean(na.rm = TRUE) %>% round(2),
        peakStartIndex = plate_data$peakStartIndex %>% min(na.rm = TRUE) %>% round(2),
        peakEndIndex = plate_data$peakEndIndex %>% max(na.rm = TRUE) %>% round(2)
      )
    )
  
  }#idx_plate
}#idx_metabolite


#merge lists together and sort

  master_list$data$peakRegionFinal <- master_list$data$peakRegionFinal %>%
    append(
      master_list$data$peakRegionSILFinal
    )

  master_list$data$peakRegionFinal <- master_list$data$peakRegionFinal[order(names(master_list$data$peakRegionFinal))]
  
```


# human check

print visuals of each peak for check solution, LTR and PQC to evaluate the peak finding alorithm

## peak Review

```{r, eval=TRUE}


#1 export plots of check and LTR and PQC

#create review folders

#create dir.

if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots"))){
    dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots"))
  }

if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/peakReview"))){
  dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/peakReview"))
}

#set plot colors
  plotColors = c(
    check = "#E41A1C" ,
    ltr = "#4DAF4A",  
    pqc = "#FF7F00"
  )

  #open pdf device
  pdf(
    onefile = TRUE, width = 11.7, height = 4,
    file = #gsub("", "-",
                paste0(
                  master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/peakReview/", Sys.Date(), "_", master_list$project_details$project_name, 
                  "_all_metabolites_peakReview.pdf"
                )
    #)
  )

  #start at metabolite level
  
  #cycle through each mrm and produce a peak review plot
  #idx_mrm = "TOL"
  i = 1
  totalMRM = length(master_list$data$peakRegionFinal %>% bind_rows() %>% arrange(metabolite) %>% .$metabolite %>% unique())
  #for(idx_mrm in c("GCA", "GCA-SIL_d4")){ # for testing
  for(idx_mrm in (master_list$data$peakRegionFinal %>% bind_rows() %>% arrange(metabolite) %>% .$metabolite %>% unique() ) ){
  
    message(paste0("printing ", i, " of ", totalMRM, " metabolites: ", idx_mrm))
    
    print(ggplot() + 
            annotate("text", x = 4, y = 25, size=8, label = idx_mrm) + 
            theme_void())

  #get plate information
  
peakReview <- list()

reviewPlotDataAll <- list()

for(idx_plate in master_list$project_details$plateList){
  print(idx_plate)
  
    #get precursor and product info from peakRegion table
  precursor_mz = master_list$data$peakRegionFinal[[idx_mrm]] %>%
    filter(sample_plate == idx_plate) %>%
    .$precursor_mz %>%
    unique()
  product_mz = master_list$data$peakRegionFinal[[idx_mrm]] %>%
    filter(sample_plate == idx_plate) %>%
    .$product_mz %>%
    unique()
  #find mrm function in the mzML file
  mzml_mrm <- master_list$data$peakRegionFinal[[idx_mrm]] %>%
    filter(sample_plate == idx_plate) %>%
    .$ms_header_id
  #print plate function 
  print(paste0("plate MS function = ", mzml_mrm))
  #peakStart in check
  peakStartRT = master_list$data$peakRegionFinal[[idx_mrm]] %>%
    filter(sample_plate == idx_plate) %>%
    .$peakStartRT %>% round(digits = 3) %>%
    unique()
  #peakEnd in check
  peakEndRT = master_list$data$peakRegionFinal[[idx_mrm]] %>%
    filter(sample_plate == idx_plate) %>%
    .$peakEndRT %>% round(digits = 3) %>%
    unique()
  #peakApex in check
  peakApexRT = master_list$data$peakRegionFinal[[idx_mrm]] %>%
    filter(sample_plate == idx_plate) %>%
    .$peakApexRT %>% round(digits = 3) %>%
    unique()

  peakReview[[idx_plate]] <- list(
    `mzML.list` = c(
      master_list$data$mzR[[idx_plate]][grepl("(?i)check", master_list$data$mzR[[idx_plate]])],
      master_list$data$mzR[[idx_plate]][grepl("(?i)_ltr", master_list$data$mzR[[idx_plate]])],
      master_list$data$mzR[[idx_plate]][grepl("(?i)_pqc", master_list$data$mzR[[idx_plate]])]
    )
  )

  #create empty list
  peakReview[[idx_plate]][["chromatograms"]] <- list()
  
  #now cycle through each mzML  
  idx_mzML = peakReview[[idx_plate]]$mzML.list[1]
  for(idx_mzML in peakReview[[idx_plate]]$mzML.list){
    
    #read in chromatogram
    
    #temp mzml object
    tempMzml <- list(
      mzR_object =  mzR::openMSfile(
        filename = paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/mzml/",idx_plate, "/", idx_mzML)
      )
    )
    
    #create mzR header
    tempMzml$mzR_header <- mzR::chromatogramHeader(tempMzml$mzR_object)
    
    #work around fix to fix issue where some projects have a method mismatch that mis-aligns the chromatogramIndex
    if(length(mzml_mrm) > 1){
      mzml_mrm = tempMzml$mzR_header %>%
        as_tibble() %>%
        filter(chromatogramIndex %in% mzml_mrm) %>%
        filter(floor(precursorIsolationWindowTargetMZ) == floor(precursor_mz)) %>%
        filter(floor(productIsolationWindowTargetMZ) == floor(product_mz)) %>%
        .$chromatogramIndex
    }

    #create mzR chromatogram
    tempMzml$mzR_chromatogram <- mzR::chromatograms(tempMzml$mzR_object)
    
    peakReview[[idx_plate]]$chromatograms[[idx_mzML]] <- tempMzml
    
  } #end idx mzML
  
  #now extract chromatogram and plot for each mzml across each plate
  #create empty data
  reviewPlotData <- list() 
  
  for(idx_mzML in names(peakReview[[idx_plate]]$chromatograms)){
    tempMzml <- list()
    tempMzml[["smooth"]] = tibble(
      time = peakReview[[idx_plate]]$chromatograms[[idx_mzML]]$mzR_chromatogram[[mzml_mrm]][["rtime"]],
      intensity = peakReview[[idx_plate]]$chromatograms[[idx_mzML]]$mzR_chromatogram[[mzml_mrm]][,2] %>%
        pracma::savgol(fl = 7, forder = 0, dorder = 0) %>%
        pracma::savgol(fl = 5, forder = 0, dorder = 0) %>%
        pracma::savgol(fl = 3, forder = 0, dorder = 0)
    ) %>% setNames(c("time", "intensity")) %>%
      add_column(.before=1,
                 sampleName = idx_mzML,
                 samplePlate = idx_plate,
                 sampleMetabolite = idx_mrm,
                 precursor_mz = precursor_mz,
                 product_mz = product_mz) %>% 
      add_column(.after = "sampleName",
                 sampleFacet = sub("(?i)_fepp_", ".", paste0(gsub("_.*", "", .$sampleName), "_", gsub(".*@", "", .$sampleName)))
      )
    
    reviewPlotData <- bind_rows(
      reviewPlotData,
      tempMzml$smooth
    )
    
  } #idx_mzML
  
  #add a sample_type column
  reviewPlotData <- reviewPlotData %>%
    add_column(sample_type = "check")
  reviewPlotData$sample_type[grepl("(?i)ltr", reviewPlotData$sampleFacet)] <- "ltr"
  reviewPlotData$sample_type[grepl("(?i)pqc", reviewPlotData$sampleFacet)] <- "pqc"
  
  
  #print plot
   print(
      ggplot(
        data = reviewPlotData, aes(x = time, y = intensity, color = sample_type)
      ) +
        geom_line(alpha = 0.5, linetype = "solid") +
        #facet_wrap(facets = "sampleFacet", scales = "free_y", ncol  = 2) +
        facet_wrap(facets = c("samplePlate", "sample_type"), scales = "free_y", ncol = 3) +
        scale_color_manual(values = plotColors) +
        geom_vline(xintercept = peakEndRT) +
        geom_vline(xintercept = peakStartRT) +
        theme_bw() +
        ggtitle(paste0(idx_mrm, "; ", idx_plate, "; integration start = ", peakStartRT, "; integration end = ", peakEndRT)) +
        scale_x_continuous(breaks = round(seq(min(reviewPlotData$time), max(reviewPlotData$time), by = 0.1), 1)) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1))
      
  )
   
  
  
  } #idx_plate

 
   i = i+1

} #idx_mrm
  
   dev.off()
  
   write_csv(
    x = master_list$data$peakRegionFinal %>% bind_rows() %>% select(sample_plate, metabolite, peakStartRT, peakEndRT) %>% arrange(metabolite, sample_plate),
    file = paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/peakReview/", Sys.Date(), "_", master_list$project_details$project_name, "_z_peakReviewTemplate.csv")
  )

```

*USER INPOUT REQUIRED*

This chunk provides the opportunity to review peak finding algorith and update the windows using the exported csv folder. To do this:

1. Navigate to ~projectID/wash[AB or N]/plots/peakReview
2. open the pdf and the csv side by side
3. scroll through the pdf and make RT updates in the CSV where needed.
4. for efficiency, only consider gross changes where clear error has occured. consider check solution, LTR and PQC

```{r}
svDialogs::dlg_message(paste0("perform peak review in folder: ", master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/peakReview/.............. and then edit template for each plate for any adjustments"))

svDialogs::dlg_message(paste0("for efficiency, only consider gross changes where clear error has occured. consider check solution, LTR and PQC"))


## read in edited template


svDialogs::dlg_message(paste0("rename the file to contain `_z_peakReviewTemplateEdited.csv` in file name"))

updateFile <- list.files(path = paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/peakReview/"), 
                         pattern = "_z_peakReviewTemplateEdited", full.names = TRUE
                         )

if(length(updateFile) > 1){
 svDialogs::dlg_message("error: too many files, there should only be 1 edited update. review and return")
  stop("too many files, there should only be 1 edited update. review and return")
}

if(length(updateFile == 1)){
  updateFile <- read_csv(updateFile)
}


#replace NA with "NA" for nicotinic acid
#updateFile$metabolite[is.na(updateFile$metabolite)] <- "NA"

for(idx_metabolite in names(master_list$data$peakRegionFinal)){
  for(idx_plate in master_list$data$peakRegionFinal[[idx_metabolite]]$sample_plate){
    #get updated times
    newStart <- updateFile %>% 
      filter(sample_plate == idx_plate & metabolite == idx_metabolite) %>% .$peakStartRT %>% as.numeric()
    # updated end
    newEnd <- updateFile %>% 
      filter(sample_plate == idx_plate & metabolite == idx_metabolite) %>% .$peakEndRT %>% as.numeric()
    #update start RT
    master_list$data$peakRegionFinal[[idx_metabolite]]$peakStartRT[which(master_list$data$peakRegionFinal[[idx_metabolite]]$sample_plate==idx_plate)] <- newStart
    #update end RT
    master_list$data$peakRegionFinal[[idx_metabolite]]$peakEndRT[which(master_list$data$peakRegionFinal[[idx_metabolite]]$sample_plate==idx_plate)] <- newEnd
     
    #update index
    #test file
    tempMzml <- list.files(path = paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/mzml/",idx_plate), full.names = TRUE)
    tempMzml <- tempMzml[grepl("(?i)_ltr", tempMzml)][1]
    tempMzml <- list(mzR_object = mzR::openMSfile(filename = tempMzml))
    #create mzR header
    tempMzml$mzR_header <- mzR::chromatogramHeader(tempMzml$mzR_object)
    #create mzR chromatogram
    tempMzml$mzR_chromatogram <- mzR::chromatograms(tempMzml$mzR_object)
    
    #set function
    mzml_mrm <- master_list$data$peakRegionFinal[[idx_metabolite]]$ms_header_id
    
    #set new index
    newStartIndex <- NA
    #set new end
    newEndIndex <- NA
    
    #update start RT
    master_list$data$peakRegionFinal[[idx_metabolite]]$peakStartIndex[which(master_list$data$peakRegionFinal[[idx_metabolite]]$sample_plate==idx_plate)] <- newStartIndex
    #update end RT
    master_list$data$peakRegionFinal[[idx_metabolite]]$peakEndIndex[which(master_list$data$peakRegionFinal[[idx_metabolite]]$sample_plate==idx_plate)] <- newEndIndex
    
    
  }#idx plate
}#idx metabolite



#master_list$data$peakRegionFinal %>% bind_rows() %>% View()

```

# peak integration

script will automatically perform peak integration

## Import mzML, extract chromData, isolate and integrate peak regions

using the identified peak regions isolate only scan data for the peak and then integrate area under the curve for each region

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#object for data for chromatogram
master_list$data$chromatogram <- list()
master_list$data$peakAreaMetabolite <- list()
master_list$data$timestamp <- list()

idx_plate = master_list$project_details$plateList[1] #for testing
#for(idx_plate in master_list$project_details$plateList[1]){ #for testing
#for(idx_plate in c("p045a")){ #for testing
for(idx_plate in master_list$project_details$plateList){

  
print(idx_plate)
  
  master_list$data$timestamp[[idx_plate]] <- list()
  
  idx_mzML = master_list$data$mzR[[idx_plate]][1] #for testing loop
  #for(idx_mzML in master_list$data$mzR[[idx_plate]][c(1:6, length(master_list$data$mzR[[idx_plate]]))]){ # for testing loop
  for(idx_mzML in master_list$data$mzR[[idx_plate]]){
    
    print(idx_mzML)
    
    #create temp mzML object
    tempMzml <- list()
    #load mzml
    tempMzml$mzR_object <- mzR::openMSfile(
      filename = paste0(master_list$project_details$plateDirList[grepl(idx_plate, master_list$project_details$plateDirList)], "/", idx_mzML)
    )
    #create mzR header
    tempMzml$mzR_header <- mzR::chromatogramHeader(tempMzml$mzR_object)
    #create mzR chromatogram
    tempMzml$mzR_chromatogram <- mzR::chromatograms(tempMzml$mzR_object)
    
    #store sample timestamp
    master_list$data$timestamp[[idx_plate]] <- bind_rows(
      master_list$data$timestamp[[idx_plate]],
      tibble(
        samplePlate = idx_plate,
        sampleName = idx_mzML,
        sampleTimestamp = tempMzml$mzR_object@backend$getRunStartTimeStamp()
      ) %>%
        add_column(.before = 1,
                   sampleID = paste0(.$samplePlate, "_", .$sampleName)
        )
    )
    
    # extract peak region
    #idx_mrm = "GCA-SIL_d4" # for testing loop
    idx_mrm ="TCA"
    idx_mrm = names(master_list$data$peakRegionFinal)[1]
    for(idx_mrm in names(master_list$data$peakRegionFinal)){
      
      #printidx_mrm)
      #for(idx_mrm in c("GCA", "GCA-SIL_d4")){ 
      #create list on first iteration of mrm
      if(!idx_mrm %in% names(master_list$data$chromatogram)){master_list$data$chromatogram[[idx_mrm]] <- list()}
      if(!idx_mrm %in% names(master_list$data$peakAreaMetabolite)){master_list$data$peakAreaMetabolite[[idx_mrm]] <- list()}
      
      #get precursor and product info from peakRegion table
      precursor_mz = master_list$data$peakRegionFinal[[idx_mrm]] %>%
        filter(sample_plate == idx_plate) %>%
        .$precursor_mz %>% unique()
      product_mz = master_list$data$peakRegionFinal[[idx_mrm]] %>%
        filter(sample_plate == idx_plate) %>%
        .$product_mz %>%
        unique()
      #find mrm function in the mzML file
      mzml_mrm <- master_list$data$peakRegionFinal[[idx_mrm]] %>%
        filter(sample_plate == idx_plate) %>%
        .$ms_header_id
      
      #work around if functions and channels have moved as a result of an RT window update
      if(length(mzml_mrm) > 1){
        
        mzml_mrm = tempMzml$mzR_header %>%
          as_tibble() %>%
          filter(chromatogramIndex %in% mzml_mrm) %>%
          filter(floor(precursorIsolationWindowTargetMZ) == floor(precursor_mz)) %>%
          filter(floor(productIsolationWindowTargetMZ) == floor(product_mz)) %>%
          .$chromatogramIndex
      }
      
      
      #peakStart in check
      peakStartRT = master_list$data$peakRegionFinal[[idx_mrm]] %>%
        filter(sample_plate == idx_plate) %>%
        filter(ms_header_id == mzml_mrm) %>%
        .$peakStartRT %>% as.numeric() %>% round(digits = 3) %>% min(na.rm = TRUE)
      #peakEnd in check
      peakEndRT = master_list$data$peakRegionFinal[[idx_mrm]] %>%
        filter(sample_plate == idx_plate) %>%
        filter(ms_header_id == mzml_mrm) %>%
        .$peakEndRT %>% as.numeric() %>% round(digits = 3) %>% max(na.rm = TRUE)
      #peakApex in check
      peakApexRT = master_list$data$peakRegionFinal[[idx_mrm]] %>%
        filter(sample_plate == idx_plate) %>%
        filter(ms_header_id == mzml_mrm) %>%
        .$peakApexRT %>% as.numeric() %>% round(digits = 3) %>% base::mean(na.rm = TRUE)
      
      
      #set empty object on first iteration
      if(!idx_plate %in% names(master_list$data$chromatogram[[idx_mrm]])){
        master_list$data$chromatogram[[idx_mrm]][[idx_plate]] <- list()
        master_list$data$chromatogram[[idx_mrm]][[idx_plate]][["all"]] <- list()
        master_list$data$chromatogram[[idx_mrm]][[idx_plate]][["peak"]] <- list()
      }
      
      
      #confirm transition is correct for the file
      updateMrm = intersect(
        which(floor(tempMzml$mzR_header$precursorIsolationWindowTargetMZ) == floor(precursor_mz)),
        which(floor(tempMzml$mzR_header$productIsolationWindowTargetMZ) == floor(product_mz))
      )
      
      #printupdateMrm)
      
      #update if necessary
      if(length(updateMrm) == 0){updateMrm = 0}
      if(!mzml_mrm %in% mzml_mrm){mzml_mrm <- updateMrm}
      
      
      #smooth chromatogram
      if(mzml_mrm > 0){
        
        #depending on mzR version header could be "rtime" or "time"
        if(length(tempMzml$mzR_chromatogram[[mzml_mrm]][["rtime"]]) > 0){mzHeader = "rtime"}
        if(length(tempMzml$mzR_chromatogram[[mzml_mrm]][["time"]]) > 0){mzHeader = "time"}
        
        #finalise chromatogram data 
        tempMzml$smooth <-  #smooth chromatogram
          tibble(
            time = tempMzml$mzR_chromatogram[[mzml_mrm]][[mzHeader]],
            intensity = tempMzml$mzR_chromatogram[[mzml_mrm]][,2] %>%
              pracma::savgol(fl = 7, forder = 0, dorder = 0) %>%
              pracma::savgol(fl = 5, forder = 0, dorder = 0) %>%
              pracma::savgol(fl = 3, forder = 0, dorder = 0)
          ) %>% 
          setNames(c("time", "intensity")) %>%
          add_column(.before=1,
                     sampleName = idx_mzML,
                     samplePlate = idx_plate,
                     sampleMetabolite = idx_mrm,
                     precursor_mz = precursor_mz,
                     product_mz = product_mz) %>% 
          add_column(.after = "sampleName",
                     sampleFacet = sub("(?i)_fepp_", ".", paste0(gsub("_.*", "", .$sampleName), "_", gsub(".*@", "", .$sampleName)))
          )
        
        
        #bind chromatogram data for all window
        master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$all <- bind_rows(
          master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$all,
          tempMzml$smooth
        )#bind rows
        
        #get peak Index
        peakStartIndex <- which.min(abs(tempMzml$smooth$time - peakStartRT))
        #set new end
        peakEndIndex <- which.min(abs(tempMzml$smooth$time - peakEndRT))
        
        #crop plot
        tempMzml$smoothCrop <- tempMzml$smooth[c(peakStartIndex:peakEndIndex),]
        
        # bind peak region
        master_list$data$chromatogram[[idx_mrm]][[idx_plate]][["peak"]] <- bind_rows(
          master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak,
          tempMzml$smoothCrop
        )
        
        
        #plot(master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$intensity)
        
        #get peak intensity using Trapz
        peakIntegral <- tempMzml$smoothCrop$intensity %>%
          pracma::trapz(.)
        
        #find peak apex
        peakApexIdx = which.min(abs(tempMzml$smooth$intensity - max(tempMzml$smoothCrop$intensity)))
        
        # find peak RT
        peakRT = tempMzml$smooth$time[peakApexIdx] %>% round(2)
        
        #combine into a master tibble
        master_list$data$peakAreaMetabolite[[idx_mrm]] <- bind_rows(
          master_list$data$peakAreaMetabolite[[idx_mrm]],
          tibble(
            samplePlate = idx_plate,
            sampleName = idx_mzML,
            sampleMetabolite = idx_mrm,
            precursor_mz = precursor_mz,
            product_mz = product_mz,
            peakAreaMetabolite = peakIntegral,
            peakRT = peakRT
          )
        )#bind_rows 
      }#if statement
      
      #create dummy data where transition doesn't exist in datafile
      if(mzml_mrm == 0){
        master_list$data$peakAreaMetabolite[[idx_mrm]] <- bind_rows(
          master_list$data$peakAreaMetabolite[[idx_mrm]],
          tibble(
            samplePlate = idx_plate,
            sampleName = idx_mzML,
            sampleMetabolite = idx_mrm,
            precursor_mz = precursor_mz,
            product_mz = product_mz,
            peakAreaMetabolite = NA,
            peakRT = NA
          )
        )
      }#if statement
      
      
    } #idx_mrm
  }#idx_mzml
}#idx_plate


#rm(tempMzml)

```

##plot chromatograms visuals

plot the chromatograms for the total project. depending on project size this can be time consuming activity. 

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#create dir.

if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots"))){
    dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots"))
  }

  if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/peakIntegrals"))){
    dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/peakIntegrals"))
  }

#object for data for chromatogram
#master_list$plots$chromatogram <- list()

#RColorBrewer::brewer.pal(n=7, name = "Set1")
#prepare colours
  plotColors = c(
    check = "#E41A1C" ,
    ltr = "#4DAF4A",  
    vltr = "#984EA3",
    sltr = "gold",
    pqc = "#FF7F00",
    sample = "#377EB8",
    blank = "darkgrey"
  )

#idx_mrm = "TRYP" # for testing loop
  i = 1
  totalMRM = length(names(master_list$data$peakRegionFinal))
for(idx_mrm in names(master_list$data$peakRegionFinal)){
   #master_list$plots$chromatogram[[idx_mrm]] <- list()
  
  message(paste0("printing plot ", i, " of ", totalMRM, "; ", idx_mrm, ".  ")) 
  
   #open png device
  pdf(onefile = TRUE, width = 16, height = 7.5,  
    file =
      paste0(
        master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/peakIntegrals/", 
        Sys.Date(), "_", master_list$project_details$project_name, "_AB_peakIntegrals_",  gsub(" ", "_", fedmatch::clean_strings(idx_mrm)), ".pdf"
    )
    )
    
   
#idx_plate="RARp003a"   
#for(idx_plate in master_list$project_details$plateList[1]){ #for testing
for(idx_plate in master_list$project_details$plateList){

  
  #print(idx_plate)

  #prepare data
  #add sampleType
  master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleType <- "sample"
  #check sample
  master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleType[grep(
    "check", master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleName, ignore.case = TRUE)] <- "check"
  #ltr sample
  master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleType[grep(
    "ltr", master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleName, ignore.case = TRUE)] <- "ltr"
  #vltr sample
  master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleType[grep(
    "vltr", master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleName, ignore.case = TRUE)] <- "vltr"
  #sltr sample
  master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleType[grep(
    "sltr", master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleName, ignore.case = TRUE)] <- "sltr"
  #pqc sample
  master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleType[grep(
    "pqc", master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleName, ignore.case = TRUE)] <- "pqc"
  #blank sample
  master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleType[grep(
    "blank", master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak$sampleName, ignore.case = TRUE)] <- "blank"
  
  #get peak start RT for vlines
  peakStart <- master_list$data$peakRegionFinal[[idx_mrm]] %>% filter(sample_plate == idx_plate) %>% .$peakStartRT %>% as.numeric() %>% min(na.rm = TRUE) %>% round(2)
  peakEnd <- master_list$data$peakRegionFinal[[idx_mrm]] %>% filter(sample_plate == idx_plate) %>% .$peakEndRT %>% as.numeric() %>%  max(na.rm = TRUE) %>% round(2)
  mrmTitle <- master_list$data$peakRegionFinal[[idx_mrm]] %>% .$mrm %>% unique()
  
  #create plot
  loopPlot <- ggplot(
    data = master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$all, 
    aes(x = time, y = intensity)) + 
    geom_vline(xintercept = peakStart, linetype = "dashed") +
    geom_vline(xintercept = peakEnd, linetype = "dashed") +
    #geom_point(shape = 4, size = 2, color = "lightgrey") +
    geom_line(na.rm = TRUE, colour = "grey") +
    geom_line(
      data = master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$peak, 
      aes(x = time, y = intensity, color = sampleType), na.rm = TRUE) +
    theme_bw() + 
    ggtitle(paste0(idx_mrm, "; ", mrmTitle, "; ", idx_plate, "; peak start = ", peakStart, "; peak end = ", peakEnd)) + 
    scale_y_continuous(labels = scales::scientific) +
    scale_colour_manual(values = plotColors) +
    facet_wrap(scales = "free_y", facets = "sampleFacet") +
    theme(
      axis.text = element_text(size = 6),
      strip.text = element_text(size = 5)
    ) +
    scale_x_continuous(breaks = round(seq(min(master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$all$time), max(master_list$data$chromatogram[[idx_mrm]][[idx_plate]]$all$time), by = 0.1), 1)) 
    
  print(loopPlot)

}#idx_plate
  
dev.off()

#iterate i
i=i+1

}#idx_mrm

#print(Sys.time() - a)


```


## CCC control charts

create quick visual control charts for the control chart check. this chunk is skipped if it is a full plate analysis (control charts happen later)

```{r}

#if(functionListMethod == "CCC"){


  
nList <- c(
  "5-HIAA", "5-HIAA-SIL_d5", "CDCA", "CDCA-SIL_d4", "GCDCA", "GCDCA-SIL_d4", "GUDCA", "GUDCA-SIL_d4", 
  "XA", "XA-SIL_d4", "12(R)-HETE", "12(R)-HETE-SIL_d8",  "15(S)-HETE","15(S)-HETE-SIL_d8", "8i-PGF2a", "8i-PGF2a-SIL_d4"
)

abList <- c(
  "3-HAA", "3-HAA-SIL_D3", "CAR(2:0)", "CAR(2:0)-SIL_d3", "CAR(4:0)", "CAR(4:0)-SIL_d3", "CAR(i5:0)", "CAR(i5:0)-SIL_d9", "KYN", "KYN-SIL_D4", "NEO", "NEO-SIL_13C5", "TRYP", "TRYP-SIL_d5", "SER", "SER-SIL_d4"
)


  #create dir.

if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots"))){
    dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots"))
  }

  if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/CCC"))){
    dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/CCC"))
  }
  
  
#make plot list
if(master_list$project_details$wash == "AB"){CCCplotList <- abList}; if(master_list$project_details$wash == "N"){CCCplotList <- nList}
CCCplotList <- CCCplotList[!grepl("-SIL", CCCplotList)] 
CCCplotList <- sort(CCCplotList)

#open PDF
pageWidth <- length(master_list$project_details$plateList)

#open pdf device
pdf(
  onefile = TRUE, width = (15 + pageWidth), height = 10,
  file = paste0(
    master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/CCC/", Sys.Date(), "_", master_list$project_details$project_name, "_SPE_", master_list$project_details$wash, "_CCC.pdf"
    
  )
)

idxPlot = CCCplotList[2]
for(idxPlot in CCCplotList){
  
  message(paste0("printing CCC for ", idxPlot)) 
  
  subPlotList <- names(master_list$data$peakAreaMetabolite)[grepl(idxPlot, names(master_list$data$peakAreaMetabolite), fixed = TRUE)]
  
  plotData <- left_join(by = "sampleName",
                        master_list$data$peakAreaMetabolite[subPlotList[!grepl("SIL", subPlotList)]][[1]] %>%
                          filter(!grepl("check", sampleName, ignore.case = TRUE)),
                        master_list$data$peakAreaMetabolite[subPlotList[grepl("SIL", subPlotList)]][[1]] %>%
                          filter(!grepl("check", sampleName, ignore.case = TRUE)) %>%
                          select(sampleName, peakAreaMetabolite) %>%
                          rename(peakAreaSil = peakAreaMetabolite)
  ) %>%
    arrange(samplePlate, sampleName) %>%
    add_column(
      peakResponse = .$peakAreaMetabolite/.$peakAreaSil,
      index = c(1:nrow(.))
    ) %>%
    select(index, samplePlate, sampleName, peakAreaMetabolite, peakAreaMetabolite, peakAreaSil, peakResponse) %>%
    pivot_longer(data = .,
                 c("peakAreaMetabolite", "peakAreaSil", "peakResponse"),
                 names_to = "dataType",
                 values_to = "value"
    ) %>%
    mutate(
      samplePlate = gsub("a", "", samplePlate),
      samplePlate = gsub("b", "", samplePlate),
      sampleType = sampleName %>% gsub(".*p0", "", .) %>% sub("_", "-", .)%>% gsub(".*-", "", .) %>% gsub("_.*", "", .)
    )
  
  #set plate boundaries
  plateBoundaryIdx = NULL
    for(idxPlate in unique(plotData$samplePlate)){
      plateBoundaryIdx <- c(
        plateBoundaryIdx, 
        plotData %>%
          filter(samplePlate == idxPlate) %>%
          .$index %>%
          max()
      )
    }
    
    plateBoundaryIdx <- plateBoundaryIdx[-length(plateBoundaryIdx)]
    
    
    #make plot

    
    plotList = list(
      title = data.frame(x = 0.5, y = 0.5, label = idxPlot) %>%
        ggplot(., aes(x = x, y = y, label = label)) +
        geom_text(size = 6) +
        theme_void() +  # removes axes and background
        xlim(0, 1) + ylim(0, 1)  # optional: set limits to centre the text
    )
    
    
    idxData = unique(plotData$dataType)[3]
    for(idxData in unique(plotData$dataType)){
    
      loopPlotData = plotData %>%
        filter(dataType == idxData)
      
      #set custom y limits
      ylimits = c(median(loopPlotData$value, na.rm = TRUE)- (IQR(loopPlotData$value, na.rm = TRUE)*4),
            median(loopPlotData$value, na.rm = TRUE) + (IQR(loopPlotData$value, na.rm = TRUE)*4)
          )
      
      #make plot
      plotList[[idxData]] <- loopPlotData %>%
        filter(sampleType != "SLTR") %>%
        ggplot(
          aes(
            x=index,
            y=value,
            fill = sampleType,
            shape = sampleType,
            size=sampleType,
            colour = sampleType
          )
        ) +
        geom_point() +
        scale_fill_manual(values = c("LTR" = "#F8766D", "PQC" = "#7CAE00", "SLTR" = "#00BFC4", "vLTR" = "black")) +
        scale_colour_manual(values = c("LTR" = "black", "PQC" = "black", "SLTR" = "#00BFC4", "vLTR" = "black")) +
        scale_shape_manual(values = c("LTR" = 21, "PQC" = 21, "SLTR" = 4, "vLTR" = 4))+
        scale_size_manual(values = c("LTR" = 4, "PQC" = 4, "SLTR" = 1.5, "vLTR" = 2)) +
        ggtitle(idxData)+ 
        theme_bw() +
        geom_vline(xintercept = plateBoundaryIdx) +
        geom_hline(yintercept = loopPlotData %>% filter(sampleType == "PQC") %>% .$value %>% median(na.rm = TRUE), linetype = "dashed", colour = "#7CAE00", linewidth = 1)+
        geom_hline(yintercept = loopPlotData %>% filter(sampleType == "LTR") %>% .$value %>% median(na.rm = TRUE), linetype = "dashed", colour = "#F8766D", linewidth = 1)+
        geom_hline(yintercept = loopPlotData %>% filter(sampleType == "vLTR") %>% .$value %>% median(na.rm = TRUE), linetype = "dashed", colour = "black", linewidth = 0.5)+
        ylim(ylimits) 
        
  }#idxData
    
    print(
      cowplot::plot_grid(
        plotlist = plotList,
        ncol = 1 ,
        rel_heights = c(0.1,1,1,1)
    )
    )
  
  } #idxPlot

  dev.off() #turn off pdf
  
  #} #if

```

#.

# checkpoint save

#.


```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/rda/"))){
  dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/rda/"))
}

save(
  master_list,
  file = paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/rda/",
                Sys.Date(), "_",
                master_list$project_details$project_name,
                "_SPE_",
                master_list$project_details$wash,
                "_checkpoint_1.rda"
  )
)

# load(
#   file = paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/rda/",
#                 Sys.Date(), "_",
#                 master_list$project_details$project_name,
#                 "_SPE_",
#                 master_list$project_details$wash,
#                 "_checkpoint_1.rda"
#   )
# )

```

# compile data

## combine data

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
#move/rename peakArea object
#master_list$data$peakAreaMetabolite <- master_list$data$peakArea

#create new object
master_list$data$peakArea <- list()

master_list$data$peakArea$Long <- left_join(by = "sampleID",
  #sample timestamp
  master_list$data$timestamp %>% bind_rows() %>% select(sampleID, sampleTimestamp),
  #peakArea
  master_list$data$peakAreaMetabolite %>% 
  bind_rows() %>%
  add_column(.after = "sampleMetabolite",
    sampleID = paste0(.$samplePlate, "_", .$sampleName)
  ) %>%
  add_column(.after = "sampleName",
    sampleType = "sample")
)
  
#check sample
master_list$data$peakArea$Long$sampleType[grep("check", master_list$data$peakArea$Long$sampleName, ignore.case = TRUE)] <- "check"
#ltr sample
master_list$data$peakArea$Long$sampleType[grep("pqc", master_list$data$peakArea$Long$sampleName, ignore.case = TRUE)] <- "pqc"
#vltr sample
master_list$data$peakArea$Long$sampleType[grep("ltr", master_list$data$peakArea$Long$sampleName, ignore.case = TRUE)] <- "ltr"
#sltr sample
master_list$data$peakArea$Long$sampleType[grep("sltr", master_list$data$peakArea$Long$sampleName, ignore.case = TRUE)] <- "sltr"
#pqc sample
master_list$data$peakArea$Long$sampleType[grep("vltr", master_list$data$peakArea$Long$sampleName, ignore.case = TRUE)] <- "vltr"
#blank sample
master_list$data$peakArea$Long$sampleType[grep("blank", master_list$data$peakArea$Long$sampleName, ignore.case = TRUE)] <- "blank"



### create mising values
#(because some transitions have noise a missing value can be recorded as non-missing)
master_list$data$peakArea$Long$peakAreaMetabolite[which(master_list$data$peakArea$Long$peakAreaMetabolite<250000)] <- NA


#make a wide version
master_list$data$peakArea$wide <- master_list$data$peakArea$Long %>%
  rename(peakArea = peakAreaMetabolite) %>%
  pivot_wider(
    id_cols = c(sampleID, sampleTimestamp, samplePlate,sampleName, sampleType),  
    names_from = sampleMetabolite, 
    values_from = peakArea
    )

#remove check, conditioning etc
master_list$data$peakArea$wide <- master_list$data$peakArea$wide %>%
  filter(sampleType == "pqc" | sampleType == "sample" | sampleType == "ltr" | sampleType == "sltr" | sampleType == "vltr") %>%
    arrange(sampleTimestamp) %>%
    add_column(sampleRunIndex = c(1:nrow(.)), .before = 1) %>%
  #add_factor column for plotting
  add_column(.after = "sampleType",
             sampleTypeFactor = factor(.$sampleType ,
                                       levels = c("sample", "pqc", "ltr", "vltr", "sltr"), 
                                       ordered = TRUE),
             sampleQC = "sample")

#select ltr as study QC type
master_list$data$peakArea$wide$sampleQC[grepl("ltr",master_list$data$peakArea$wide$sampleType)] <- "qc"
master_list$data$peakArea$wide$sampleQC[grepl("sltr",master_list$data$peakArea$wide$sampleType)] <- "sample"
master_list$data$peakArea$wide$sampleQC[grepl("vltr",master_list$data$peakArea$wide$sampleType)] <- "sample"


#remove neg values
tempData <-  master_list$data$peakArea$wide %>% select(!contains("sample")) %>% as.matrix()
tempData[which(tempData <0)] <- 0
master_list$data$peakArea$wide <- bind_cols(
  master_list$data$peakArea$wide %>% select(contains("sample")),
  tempData %>% as_tibble()
)



#remove sltr (only x per plate)
master_list$data$peakArea$wide <- master_list$data$peakArea$wide %>% 
  filter(sampleType != "sltr")

```

## Filter failed samples

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#first flag samples where addition of SIL may have failed

silSigal <- tibble(
  sampleID = master_list$data$peakArea$wide %>%
    .$sampleID,
  summedSignal = master_list$data$peakArea$wide %>%
    select(!contains("sample")) %>%
    select(contains("-sil")) %>%
    rowSums(na.rm = TRUE)
)

metaboliteSignal <- tibble(
  sampleID = master_list$data$peakArea$wide %>%
    .$sampleID,
  summedSignal = master_list$data$peakArea$wide %>%
    select(!contains("sample")) %>%
    select(!contains("-sil")) %>%
    rowSums(na.rm = TRUE)
)


flaggedSamples <- unique(
  c(
    silSigal[["sampleID"]][which(silSigal$summedSignal < (median(silSigal$summedSignal)*0.01))],
  metaboliteSignal[["sampleID"]][which(metaboliteSignal$summedSignal < (median(metaboliteSignal$summedSignal)*0.01))]
  ))

master_list$data$peakArea$wideFiltered <- master_list$data$peakArea$wide %>%
  filter(!sampleID %in% flaggedSamples)

```

## Filter Failed features (missing value)

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

flaggedFeature <- NULL

#idx_metabolite <- "8-OH-2-dG"
#idx_metabolite = names(master_list$data$peakArea$wide %>% select(!contains("sample")))[1] #for testign loop
for(idx_metabolite in names(master_list$data$peakArea$wide %>% select(!contains("sample")))){
  #calculate for sample
  sampleMissing <- which(is.na(master_list$data$peakArea$wideFiltered %>% filter(sampleType == "sample") %>% .[[idx_metabolite]])) %>% length()
  samplePercent <- (100/(master_list$data$peakArea$wideFiltered %>% filter(sampleType == "sample") %>% nrow()))*sampleMissing
  #calculate for pqc
  pqcMissing <- which(is.na(master_list$data$peakArea$wideFiltered %>% filter(sampleType == "pqc") %>% .[[idx_metabolite]])) %>% length()
  pqcPercent <- (100/(master_list$data$peakArea$wideFiltered %>% filter(sampleType == "pqc") %>% nrow()))*pqcMissing
  #caculcate for ltr
  ltrMissing <- which(is.na(master_list$data$peakArea$wideFiltered %>% filter(sampleType == "ltr") %>% .[[idx_metabolite]])) %>% length()
  ltrPercent <- (100/(master_list$data$peakArea$wideFiltered %>% filter(sampleType == "ltr") %>% nrow()))*ltrMissing
  
  #if is a non-SIL standard
  if(!grepl("SIL", idx_metabolite)){
    if(samplePercent > 75.5 | pqcPercent > 25.5){ #} | ltrPercent >20){
      flaggedFeature = c(flaggedFeature, idx_metabolite)
    }
  }
  
  #if it is a SIL standard
  if(grepl("SIL", idx_metabolite)){
    if(samplePercent > 15.5 | pqcPercent > 15.5){ # | ltrPercent >25){
      flaggedFeature = c(flaggedFeature, idx_metabolite)
    }
  }
  
}

#create final dataSet
master_list$data$peakArea$wideFiltered <- master_list$data$peakArea$wideFiltered %>%
  select(!all_of(flaggedFeature))

```

## record metabolite metadata

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#keep a record of the metabolite names before and after clean names
master_list$templates[["metaboliteNames"]] <- tibble(
  oldNames = master_list$data$peakArea$wideFiltered %>% names(),
  newNames = master_list$data$peakArea$wideFiltered %>% clean_names() %>% names()
) %>%
  add_column(
    tidyName = gsub("\\..*", "", gsub("-QC", "", gsub("_.*", "", .$oldNames))) %>% toupper(),
    experiment = NA,
    polarity = NA,
    rt = NA,
    silUsed = NA#,
    #pqcRSD = NA
  ) %>%
  filter(!grepl("sample", oldNames))

#add full name from scout master
idx_name = (master_list$templates[["metaboliteNames"]][["tidyName"]])[1]
for(idx_name in (master_list$templates[["metaboliteNames"]][["tidyName"]])){
  
  oldName = master_list$templates[["metaboliteNames"]]$oldNames[which(master_list$templates[["metaboliteNames"]]$tidyName == idx_name)]
  
  #add experiment type
  experiment = master_list$templates$msMethod$experiment[which(toupper(master_list$templates$msMethod$metabolite) == toupper(oldName))] %>% toupper() %>% unique()
  #add to meta table
     if(length(experiment)==1){
  master_list$templates[["metaboliteNames"]]$experiment[which(master_list$templates[["metaboliteNames"]]$oldNames == oldName)] <-
   experiment
  }
  
  #find polarity
  polarity = master_list$templates$msMethod$polarity[which(toupper(master_list$templates$msMethod$metabolite) == toupper(oldName))] %>% tolower() %>% unique()
  
  #add polarity to meta table
   if(length(polarity)==1){
  master_list$templates[["metaboliteNames"]]$polarity[which(master_list$templates[["metaboliteNames"]]$oldNames == oldName)] <-
   polarity
  }

  #find rt
  rt = mean(na.rm=TRUE,
    c(
    master_list$data$peakRegionFinal[[oldName]] %>% .$peakStartRT %>% as.numeric(), 
    master_list$data$peakRegionFinal[[oldName]] %>% .$peakEndRT %>% as.numeric())
  ) %>% round(3)
 
  #add RT to meta table
  master_list$templates[["metaboliteNames"]]$rt[which(master_list$templates[["metaboliteNames"]]$oldNames == oldName)] <- rt
  
  }#close idx_name loop



```

## Sorted

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#split into sub-plates
master_list$data$peakArea[["sorted"]] <- list()
for(idx_plate in unique(master_list$data$peakArea$wideFiltered$samplePlate)){
  master_list$data$peakArea$sorted[[idx_plate]] <- master_list$data$peakArea$wideFiltered %>% filter(samplePlate == idx_plate) %>% clean_names() 
}

```

## Impute

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

## 1.4. impute missing values [min/2 imputation (missing assumed < LOD)] -----------------------------------------------------
#Imputation of the all zero value and missing data 
#Imputation is completed using x/2, where x is minimum intensity of that feature in the batch

#impute function
#add LGW impute function
lgw_impute <- function(x){
  map(.x = x, .f = ~ (min(.x[.x > 0], na.rm = TRUE))/2) %>%
    #use replace_na to replace NAs with min/2 value
    replace_na(
      data = x %>% mutate_all(~ replace(., . == 0, NA)), #note - replace zeros with NA to make compatible with replace_na()
      replace = .) #note - replace with list of min/2 values generated from map function in pipe (.)
}

#set data list
master_list$data$peakArea$imputed <- list()
#run loop in all plates
for(idx_plate in names(master_list$data$peakArea$sorted)){
  #set all 0, NaN, is.infinate to a NA value for consistency
  #create matrix 
  master_list$data$peakArea$imputed[[idx_plate]]  <- master_list$data$peakArea$sorted[[idx_plate]] %>%
    column_to_rownames("sample_id") %>%
    select(-contains("sample")) %>%
    as.matrix()
  
  #replace 0, NaN and Inf with NA for imputation (min/2)
  master_list$data$peakArea$imputed[[idx_plate]][master_list$data$peakArea$imputed[[idx_plate]] ==0] <- NA
  master_list$data$peakArea$imputed[[idx_plate]][is.infinite(master_list$data$peakArea$imputed[[idx_plate]])] <- NA
  master_list$data$peakArea$imputed[[idx_plate]][is.nan(master_list$data$peakArea$imputed[[idx_plate]])] <- NA
  
  
  #run lgw_impute function
  master_list$data$peakArea$imputed[[idx_plate]] <- master_list$data$peakArea$imputed[[idx_plate]] %>% 
    as.data.frame() %>%
    lgw_impute() %>%
    rownames_to_column("sample_id") %>%
    as_tibble() %>%
    #samples that are 100% missing (zero/na) have imputation error as min/2(x) is inf. So this step replaces inf with zeros (will be flagged for filter later anyway).
    mutate_all(function(x) ifelse(is.infinite(x), 1, x)) %>%
    left_join(
      master_list$data$peakArea$sorted[[idx_plate]] %>%
        select(contains("sample")),
      .,
      by = "sample_id"
    ) %>%
    add_column(
      sampleDataSource = ".peakArea"
    ) %>% clean_names()
}

```


## combat raw signal

perform combat batch correction across plates in event of plate to plate variation

```{r}

# combat takes into account all features in the algorithm. Because internal standards and samples are added at different stages their batch correction needs may differ. I have factored this into the next steps

master_list$data$peakArea$combat <- list()

# for SIL
dataImputed <- master_list$data$peakArea$imputed %>% 
  bind_rows() %>%
  select(
    contains("sample"),
    contains("_sil", ignore.case = TRUE)
  )
#data prep (samples in cols and metabolites in rows)

dataPreCombat <- dataImputed %>%
  select(sample_id, !contains("sample")) %>%
  #offset by constant factor of 1 for lof step later
   mutate(
    across(where(is.numeric), ~ .x+1)
    ) %>%
  column_to_rownames("sample_id") %>%
  t() %>%
  log10()

#extract batch
batch = dataImputed %>%
  .$sample_plate

#perform batch correction
dataPostCombat <- sva::ComBat(
  dat = as.matrix(dataPreCombat),
  batch = batch,
  mod = NULL,           # or NULL
  par.prior = TRUE,
  prior.plots = FALSE,
  mean.only = FALSE          # set TRUE if you only want mean-shift correction
  #ref.batch = 1
)

# sometimes combat gives neg values, offset factor constant to remove this option
for(idx_metaboliteRow in 1:nrow(dataPostCombat)){
  minRow = min(dataPostCombat[idx_metaboliteRow,], na.rm = TRUE)
  if(minRow < 0){
    dataPostCombat[idx_metaboliteRow, ] <- dataPostCombat[idx_metaboliteRow, ] + abs(minRow*2)   
  }
}

# unlog data
dataPostCombat2 <- left_join(
  by = "sample_id",
  dataImputed %>% select(contains("sample")),
  ((10^dataPostCombat) - 1) %>% # per-feature log nd offset removal
    round(3) %>%
    t(.) %>%
    as.data.frame() %>%
    lgw_impute() %>%
    rownames_to_column("sample_id") %>%
    as_tibble()
)


# repeat for non Sil data
# for SIL
dataImputed <- master_list$data$peakArea$imputed %>% 
  bind_rows() %>%
  select(
    contains("sample"),
    !contains("_sil", ignore.case = TRUE)
  )
#data prep (samples in cols and metabolites in rows)

dataPreCombat <- dataImputed %>%
  select(sample_id, !contains("sample")) %>%
  #offset by constant factor of 1 for lof step later
   mutate(
    across(where(is.numeric), ~ .x+1)
    ) %>%
  column_to_rownames("sample_id") %>%
  t() %>%
  log10()

#extract batch
batch = dataImputed %>%
  .$sample_plate

#perform batch correction
dataPostCombat <- sva::ComBat(
  dat = as.matrix(dataPreCombat),
  batch = batch,
  mod = NULL,           # or NULL
  par.prior = TRUE,
  prior.plots = FALSE,
  mean.only = FALSE          # set TRUE if you only want mean-shift correction
  #ref.batch = 1
)

# sometimes combat gives neg values, offset factor constant to remove this option
for(idx_metaboliteRow in 1:nrow(dataPostCombat)){
  minRow = min(dataPostCombat[idx_metaboliteRow,], na.rm = TRUE)
  if(minRow < 0){
    dataPostCombat[idx_metaboliteRow, ] <- dataPostCombat[idx_metaboliteRow, ] + abs(minRow*2)   
  }
}

# unlog data
dataPostCombat2 <- dataPostCombat2 %>%
  left_join(
  by = "sample_id",
  .,
  ((10^dataPostCombat) - 1) %>% # per-feature log nd offset removal
    round(3) %>%
    t(.) %>%
    as.data.frame() %>%
    lgw_impute() %>%
    rownames_to_column("sample_id") %>%
    as_tibble()
) %>%
  select(
    master_list$data$peakArea$imputed %>% 
      bind_rows() %>%
      names()
  )


#return to per plate in master data

 
for(idx_plate in master_list$data$peakArea$impute %>% names()){
  master_list$data$peakArea$combat[[idx_plate]] <- dataPostCombat2 %>%
    filter(sample_plate == idx_plate)
}


```


## statTarget raw signal

```{r}

## 1.5. statTarget signalDrift | batch correction ----------------------------

#create batch correction directory

skylineR_directory <- paste0(master_list$project_details$project_dir, "\\", master_list$project_details$wash)

#data

if(!dir.exists(paste0(skylineR_directory, "\\statTarget"))){
  dir.create(paste0(skylineR_directory, "\\statTarget"))
}


#set qc-type
#set dummy value for while loop
master_list$project_details$statTarget_qc_type <- "hold"

#run while loop
while(master_list$project_details$statTarget_qc_type != "LTR" & master_list$project_details$statTarget_qc_type != "VLTR" & master_list$project_details$statTarget_qc_type != "PQC"){

master_list$project_details$statTarget_qc_type <- dlgInput("which qc type will be used for statTarget", "VLTR/LTR/PQC - default is PQC")$res

}

#create data list 
FUNC_list <- list()
FUNC_list$project_dir <- paste0(skylineR_directory,
                                "\\statTarget")

#set up project folders for batch correct
#create directories 
if(!dir.exists(paste0(FUNC_list$project_dir,"/", Sys.Date(), "_sigCorResults"))){
  dir.create(paste0(FUNC_list$project_dir,"/", Sys.Date(), "_sigCorResults"))
}

setwd(paste0(FUNC_list$project_dir,"/", Sys.Date(), "_sigCorResults")) 

#apply on peakArea data (post-impute)
#set master data for function
FUNC_list$master_data <- bind_rows(master_list$data$peakArea$imputed)
#set_qc type used for signal drift correction
FUNC_list$master_data[["sample_type"]] <- "sample"
FUNC_list$master_data[["sample_type"]][which(tolower(FUNC_list$master_data[["sample_type_factor"]]) == tolower(master_list$project_details$statTarget_qc_type))] <- "qc"

#flag QCs that have failed because of mis-injection (very low signal (<10 % of median))
temp.rowSums <- FUNC_list$master_data %>%
  filter(sample_type_factor == master_list$project_details$statTarget_qc_type) %>%
  select(!contains("sample")) %>%
  rowSums()
#find file names
temp.qcFail <- (FUNC_list$master_data  %>%
                  filter(sample_type_factor == master_list$project_details$statTarget_qc_type) %>%
                  .$sample_name)[which(temp.rowSums < median(temp.rowSums*0.1))]

#reset failed QC injections to "sample" so is not included in statTarget algorithm
FUNC_list$master_data[["sample_type"]][which(FUNC_list$master_data[["sample_name"]] %in% temp.qcFail)] <- "sample"


#flag low number of QCs using ratio of QC to total samples
for(idxPlate in names(master_list$data$peakArea$imputed)){
  
  plateData <- FUNC_list$master_data %>% filter(sample_plate == idxPlate)
  totalSamples <- length(plateData %>% .$sample_type_factor)
  requiredQCs <- (6/96) # Minimum of 6 QC for 96 samples
  qcCount <- length(which(tolower(plateData$sample_type) == "qc"))
  qcCountRatio <- (qcCount/totalSamples)
  dlg_message(paste0(idxPlate, " has ", qcCount, " ", master_list$project_details$statTarget_qc_type, "s for use by statTarget"), "ok")
  
  if(qcCountRatio < requiredQCs || qcCount < 2){
    dlg_message(paste0("You do not have enough QCs for statTarget on plate ",idxPlate,". Stopping qcCheckeR. Please remove plate ",idxPlate," and re-run chunk."), type = 'ok')
    FUNC_list$corrected_data$data <- NULL
  }
}


#exclude failed qcs (see chunk above to id failed QCs)
FUNC_list$master_data[["sample_type"]]


#set metabolite list
FUNC_list$metabolite_list <- master_list$data$peakArea$imputed %>%
  bind_rows() %>%
  select(-contains("sample")) %>%
  names()

#### 1.5.a. create the required metadata file (PhenoFile) for statTarget::shiftCor---------------------------------------------------------------
FUNC_list$PhenoFile <- list()
# build PhenoFile file template
FUNC_list$PhenoFile$template <- FUNC_list$master_data %>% 
  select(all_of("sample_name")) %>%
  dplyr::rename(sample = all_of("sample_name")) %>%
  add_column(FUNC_list$master_data %>% 
               select(all_of("sample_name"))) %>%
  add_column(FUNC_list$master_data %>%
               select(all_of("sample_plate"))) %>%
  add_column(FUNC_list$master_data %>%
               select(all_of("sample_type"))) %>%
  dplyr::rename(class = all_of("sample_type")) %>% 
  add_column(FUNC_list$master_data %>%
               select(all_of("sample_type"))) %>%
  add_column(FUNC_list$master_data %>%
               select(all_of("sample_run_index")))

#### arrange by run order
FUNC_list$PhenoFile$template <- FUNC_list$PhenoFile$template %>%
  arrange_at("sample_run_index")

#### QC placement
#stat target needs a qc at postion 1 and last position in the run. Default run order takes this into account, but some datasets this is not the case
#in this instance this section of code artificially moves the first and last qc into position. It is completed for each batch.

FUNC_list$PhenoFile$template_qc_order <- NULL
qc_idx <- NULL
for(idx_batch in FUNC_list$PhenoFile$template %>% 
    select(all_of("sample_plate")) %>%
    unique() %>%
    as.matrix() %>%
    c()
){
  
  #create a temp tibble batch specific
  loop_temp_data <- FUNC_list$PhenoFile$template %>%
    filter(!!as.symbol("sample_plate") == idx_batch)
  
  #ensure a sample_type "qc" is "first" and "last" in the worklist order. Required for statTarget::shiftCor
  loop_qc_idx <- which(loop_temp_data %>% 
                         select(all_of("sample_type")) 
                       == "qc")
  # browser()
  #if qc is not run before the samples - artificially move first qc to run order position 1. This is required for statTarget
  if(loop_qc_idx[1] > 1){
    loop_temp_data <- loop_temp_data %>%
      slice(loop_qc_idx[1],1:nrow(loop_temp_data)) %>%
      slice(-(loop_qc_idx[1]+1))
  }
  
  #create last qc
  if(loop_qc_idx[length(loop_qc_idx)] < nrow(loop_temp_data)){
    loop_temp_data <- loop_temp_data %>%
      slice(1:nrow(loop_temp_data), loop_qc_idx[length(loop_qc_idx)]) %>%
      slice(-loop_qc_idx[length(loop_qc_idx)])
  }
  
  #create total qc_idx for use later
  qc_idx <- c(qc_idx,
              loop_qc_idx)
  
  FUNC_list$PhenoFile$template_qc_order <- bind_rows(FUNC_list$PhenoFile$template_qc_order,
                                                     loop_temp_data)
  
}

#set sample column for statTarget requires "QC" in QC rows, and sample name in sample rows
FUNC_list$PhenoFile$template_qc_order$sample[which(FUNC_list$PhenoFile$template_qc_order %>% 
                                                     select(all_of("sample_type")) == "qc")] <- paste0("QC", rep(1:length(qc_idx)))
FUNC_list$PhenoFile$template_qc_order$sample[which(FUNC_list$PhenoFile$template_qc_order %>% 
                                                     select(all_of("sample_type")) == "sample")] <- paste0("sample", 
                                                                                                           rep(1:(nrow(FUNC_list$PhenoFile$template_qc_order)-length(qc_idx))))
#set NA for class column in rows that are NA
FUNC_list$PhenoFile$template_qc_order$class[which(FUNC_list$PhenoFile$template_qc_order %>% 
                                                    select(all_of("sample_type")) == "qc")] <- NA

#rename column header for statTarget template
FUNC_list$PhenoFile$template_sample_id <- FUNC_list$PhenoFile$template_qc_order %>% 
  dplyr::rename(sample_id = all_of("sample_name"),
                batch = all_of("sample_plate"),
                order = all_of("sample_run_index")) %>%
  select(sample, batch, class, order, sample_id)

#confirm order columnn is continuous
FUNC_list$PhenoFile$template_sample_id$order <- c(1:nrow(FUNC_list$PhenoFile$template_sample_id))

#set batch/plate - numeric value starting at 1 - max number of plates/batch
temp_batch <- 1
for(idx_batch_set in unique(FUNC_list$PhenoFile$template_sample_id$batch)){
  FUNC_list$PhenoFile$template_sample_id$batch[which(FUNC_list$PhenoFile$template_sample_id$batch == idx_batch_set)] <- temp_batch
  temp_batch <- temp_batch + 1
}

FUNC_list$PhenoFile$template_sample_id$batch <- FUNC_list$PhenoFile$template_sample_id$batch %>%
  as.numeric()

#final Phenofile
FUNC_list$PhenoFile$PhenoFileOut <- FUNC_list$PhenoFile$template_sample_id %>%
  select(-sample_id)

# write out as csv (requirement for statTarget::shiftCor)
write_csv(x = FUNC_list$PhenoFile$PhenoFileOut,
          file = paste(getwd(), "/PhenoFile.csv", sep="")
)

#### 1.5.b. create data for statTarget::shiftCor  -----------------------------------
FUNC_list$ProfileFile <- list()

#must have samples in columns and metabolites in rows
FUNC_list$ProfileFile$template  <- FUNC_list$master_data %>%
  select(all_of("sample_name"),
         all_of(FUNC_list$metabolite_list)) %>%
  dplyr::rename(sample_id = !!"sample_name")

#match run order to PhenoFile
FUNC_list$ProfileFile$template_qc_order <- FUNC_list$PhenoFile$template_sample_id %>%
  select(sample, sample_id) %>%
  left_join(FUNC_list$ProfileFile$template, by = "sample_id") %>%
  select(-sample_id)

#transpose tibble for statTarget
FUNC_list$ProfileFile$ProfileFile <- as_tibble(
  cbind(nms = names(FUNC_list$ProfileFile$template_qc_order), 
        t(FUNC_list$ProfileFile$template_qc_order))
) %>%
  setNames(.[1,]) %>%
  dplyr::rename(name = sample) %>%
  filter(name != "sample") %>%
  mutate(across(!contains("name", ignore.case = FALSE), as.numeric))

#create a metabolite list and create metabolite code
FUNC_list$ProfileFile$metabolite_list <- FUNC_list$ProfileFile$ProfileFile %>%
  select(name) %>%
  add_column(metabolite_code = paste0("M", rep(1:nrow(FUNC_list$ProfileFile$ProfileFile))))

#add metabolite code to data
FUNC_list$ProfileFile$ProfileFile <- left_join(
  FUNC_list$ProfileFile$metabolite_list,
  FUNC_list$ProfileFile$ProfileFile,
  by = "name") %>%
  select(-name) %>%
  dplyr::rename(name = metabolite_code)

# write out as csv (requirement for statTarget::shiftCor)
write_csv(x = FUNC_list$ProfileFile$ProfileFile, 
          file = paste0(getwd(), "/ProfileFile.csv")
)

#script files
samPeno <- paste(getwd(), "\\PhenoFile.csv", sep="")
samFile <- paste(getwd(),  "\\ProfileFile.csv", sep="")

#create_directories
if(!dir.exists(paste(getwd(), "shiftCor", sep = "\\"))){
  dir.create(paste(getwd(), "shiftCor", sep = "\\"))
}

#and subDir
if(!dir.exists(paste(getwd(), "shiftCor", "Before_shiftCor", sep = "\\"))){
  dir.create(paste(getwd(), "shiftCor", "Before_shiftCor", sep = "\\"))
}


#check nchar
if(nchar(paste(getwd(), "shiftCor", "Before_shiftCor", sep = "\\")) >= 225){
  svDialogs::dlg_message(paste0("WARNING: your path length is ", nchar(paste(getwd(), "shiftCor", "Before_shiftCor", sep = "\\")), " characters. Beware - windows path limit is 260, you may experience a path error. If chunk errors, relocate project to a shorter path (e.g. desktop) for processing"))
}

                 
#run statTarget
statTarget::shiftCor(samPeno = samPeno,
                     samFile =  samFile,
                     Frule = 0,
                     ntree = 500,
                     MLmethod = 'QCRFSC',
                     imputeM = "minHalf",
                     plot = FALSE,
                     coCV = 10000
)

#read back in csv

FUNC_list$corrected_data$data <- read_csv(
  paste0(getwd(), "/statTarget/shiftCor/After_shiftCor/shift_all_cor.csv"),
  show_col_types = FALSE)

#different versions of r/statTarget appear to produce shift_all_cor in different orientations. Some have metabolites in columns some in rows. Unsure as to why, I did not find out which package inside statTarget was producing this outcome.
#the next phase fixes this by performing a check as to the orientation of the csv

#older packages (metab lites in rows)
if("sample1" %in% colnames(FUNC_list$corrected_data$data)){
  FUNC_list$corrected_data$data  <- FUNC_list$corrected_data$data  %>%
    filter(sample != "class") %>%
    dplyr::rename(name = sample) %>%
    mutate(across(!contains("name", ignore.case = FALSE), as.numeric))
}

#updated packages (metabolites in columns)
if("M1" %in% colnames(FUNC_list$corrected_data$data)){
  FUNC_list$corrected_data$data  <- FUNC_list$corrected_data$data  %>%
    t() %>% data.frame() %>%
    rownames_to_column() %>% 
    as_tibble() %>%
    setNames(.[1,]) %>%
    filter(sample != "class" & sample != "sample" ) %>%
    dplyr::rename(name = sample) %>%
    mutate(across(!contains("name", ignore.case = FALSE), as.numeric))
}

#recombine with sample filenames and lipid names
FUNC_list$corrected_data$data_transposed <- right_join(
  FUNC_list$ProfileFile$metabolite_list %>% dplyr::rename(lipid = metabolite_code),
  FUNC_list$corrected_data$data %>% dplyr::rename(lipid = name),
  by = "lipid"
) %>%
  select(-lipid) %>%
  as.matrix() %>%
  t() %>% data.frame() %>%
  rownames_to_column() %>% 
  as_tibble() %>%
  setNames(.[1,]) %>%
  filter(name!="name") %>%
  mutate(across(!contains("name", ignore.case = FALSE), as.numeric)) %>%
  dplyr::rename(sample = name) %>%
  left_join(x = FUNC_list$PhenoFile$template_sample_id,
            y = .,
            by = "sample") %>%
  dplyr::rename(!!"sample_name" := sample_id) %>%
  left_join(
    x = FUNC_list$master_data %>%
      select(contains("sample")),
    y = .,
    by = "sample_name"
  ) %>%
  select(-all_of(c("sample", "batch", "class", "order")))

#### 1.5.c. post-statTarget peak area mean adjustment ------------------------------------
#because the statTarget correction changes the output signal area of the lipids, this next section re-scales the values based on the change (ratio) between pre and post corrected signal mean in the QCs
#step one - get mean value for each metabolite in the QC samples - pre-single drift corrected data 
FUNC_list$corrected_data$qc_means <- FUNC_list$master_data %>%
  filter(!!as.symbol("sample_type") == "qc") %>%
  select(-contains("sample")) %>%
  colMeans() %>%
  as_tibble() %>%
  dplyr::rename(original_mean = value) %>%
  add_column(metabolite = FUNC_list$master_data %>%
               select(-contains("sample")) %>%
               names(), 
             .before = "original_mean") %>%
  #step two - get mean value for each metabolite in the QC samples - post-single drift corrected data 
  left_join(.,
            FUNC_list$corrected_data$data_transposed %>%
              filter(!!as.symbol("sample_type") == "qc") %>%
              select(-contains("sample")) %>%
              colMeans() %>%
              data.frame %>%
              rownames_to_column() %>%
              as_tibble() %>%
              setNames(c("metabolite", "corrected_mean")),
            by = "metabolite") %>%
  #step three - create ratio factor for concentration adjustment
  add_column(correction_ratio = .$corrected_mean/.$original_mean)

#step 4 - adjust data concentrations
FUNC_list$corrected_data$data_qc_mean_adjusted <- FUNC_list$corrected_data$data_transposed
#run loop
for(idx_metabolite in FUNC_list$corrected_data$qc_means$metabolite){
  FUNC_list$corrected_data$data_qc_mean_adjusted[[idx_metabolite]] <- FUNC_list$corrected_data$data_qc_mean_adjusted[[idx_metabolite]]/
    FUNC_list$corrected_data$qc_means[["correction_ratio"]][which(FUNC_list$corrected_data$qc_means[["metabolite"]]==idx_metabolite)]
}

#return QC type to ltr data
FUNC_DATA_qc_type <- unique(bind_rows(master_list$data$peakArea$sorted)$sample_type_factor[which(bind_rows(master_list$data$peakArea$sorted)$sample_type =="qc")]) %>% as.character()
FUNC_list$corrected_data$data_qc_mean_adjusted$sample_type <- FUNC_list$corrected_data$data_qc_mean_adjusted$sample_type_factor
#FUNC_list$corrected_data$data_qc_mean_adjusted$sample_type[which(FUNC_list$corrected_data$data_qc_mean_adjusted$sample_type_factor == FUNC_DATA_qc_type)] <- "qc"

FUNC_list

#list for storing signal drift corrected data (per project)
master_list$data$peakArea$statTarget <- list()
#loop to back into plate elements of list
for(idx_batch in unique(FUNC_list$corrected_data$data_qc_mean_adjusted$sample_plate)){
  master_list$data$peakArea$statTarget[[idx_batch]] <- FUNC_list$corrected_data$data_qc_mean_adjusted %>%
    filter(sample_plate == idx_batch) %>%
    mutate(across(where(is.numeric), round, 0))
  #set dataSource
  master_list$data$peakArea$statTarget[[idx_batch]]$sample_data_source <- "peakArea.statTarget"
}

#reset wd to project dir
setwd(master_list$project_details$project_dir)

#remove extra files
rm(list = c(ls()[which(ls() != "master_list")]))


```



# peakResponse

## for matched internal standards

```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

#set up objects lists for data storage
master_list$data$response <- list()

#reset table
master_list$templates$metaboliteNames$matchedSilUsed <- NA; master_list$templates$metaboliteNames$silUsed_imputed <- NA; master_list$templates$metaboliteNames$silUsed_combat <- NA; master_list$templates$metaboliteNames$silUsed_statTarget <- NA; master_list$templates$metaboliteNames$silUsed <- NA; master_list$templates$metaboliteNames



#master_list$templates$metaboliteNames[["silUsed"]] = NA

idx_dataType = "imputed"
for(idx_dataType in c("imputed", "combat", "statTarget")){
  
  master_list$data$response[[idx_dataType]] <- list()
  
  #first calculate response for matched internal standards
  silList <- master_list$data$peakArea[[idx_dataType]] %>% 
    bind_rows() %>%
    select(contains("_sil")) %>%
    names()
  
  #loopData 
  loopData <- master_list$data$peakArea[[idx_dataType]] %>%
    bind_rows()
  
  idx_sil = silList[15]
  for(idx_sil in silList){
    #print(idx_sil)
    
    #find target
    loopTarget <- gsub("_sil.*", "", idx_sil) 
    
   
    
    if(length(which(names(loopData) == loopTarget)) == 1){
      
      #store SIL used  
      master_list$templates$metaboliteNames[[paste0("silUsed_", idx_dataType)]][which(master_list$templates$metaboliteNames$newNames == loopTarget)] <- 
        master_list$templates$metaboliteNames$oldNames[which(master_list$templates$metaboliteNames$newNames ==  idx_sil)]
      
      
      #turn matchedSilused to TRUE
      master_list$templates$metaboliteNames$matchedSilUsed[which(master_list$templates$metaboliteNames$newNames == loopTarget)] <- 
        TRUE
      

  for(idx_plate in names(master_list$data$peakArea[[idx_dataType]])){
    
    #create dataframe for first iteration of loop
    if(is.null(master_list$data$response[[idx_dataType]][[idx_plate]])){
    
    master_list$data$response[[idx_dataType]][[idx_plate]] <- master_list$data$peakArea[[idx_dataType]][[idx_plate]] %>% 
      select(contains("sample")) 
    
    #change data source
    master_list$data$response[[idx_dataType]][[idx_plate]]$sample_data_source <- "imputeResponse"
    
    
    }# is.null
    
    
    #set loop data
    loopData <- master_list$data$peakArea[[idx_dataType]][[idx_plate]]
    
    #get metabolte target data
    loopTargetData <- loopData[[loopTarget]]
    
    #getSil data
    loopSilData <- loopData[[idx_sil]]
    
    #create response
    loopResponse <- tibble(
      loopResponse = loopTargetData/loopSilData
    ) %>% rename_with(~loopTarget)
    
    #store data
    master_list$data$response[[idx_dataType]][[idx_plate]] <- master_list$data$response[[idx_dataType]][[idx_plate]] %>%
      bind_cols(
        loopResponse
      )
        
        
      } #idxPlate
      
    } #idx_sil
    
  } #idx_dataType
  
} #if
  

master_list$templates$metaboliteNames


```

### for surrogate internal standards

```{r}

#get list of metabolites
metaboliteList <- master_list$templates$metaboliteNames[-which(master_list$templates$metaboliteNames$matchedSilUsed == TRUE),] %>%
  filter(!grepl("sil", newNames, ignore.case = TRUE)) %>% 
  .$newNames


  
idx_dataType = "imputed"
for(idx_dataType in c("imputed", "combat", "statTarget")){
  
  #print(idx_dataType)
  

  #set loop data
  loopData <- master_list$data$peakArea[[idx_dataType]] %>% bind_rows()
  loopDataQC <- loopData %>% filter(sample_type == "pqc")
  
  #find best surrogate based on PQC rsd
  idx_metabolite = metaboliteList[1]
  for(idx_metabolite in metaboliteList){
    #print(idx_metabolite)
    #get experiment
    metaboliteExperiment <- master_list$templates$metaboliteNames$experiment[which(master_list$templates$metaboliteNames$newNames == idx_metabolite)]
    #get polarity
    metabolitePolarity <- master_list$templates$metaboliteNames$polarity[which(master_list$templates$metaboliteNames$newNames == idx_metabolite)]
    #getRT
    metaboliteRT <- master_list$templates$metaboliteNames$rt[which(master_list$templates$metaboliteNames$newNames == idx_metabolite)]
    
    
    #get loop qc data per metabolite
    loopMetaboliteData <- loopDataQC[[idx_metabolite]]
    
    #loop through sil using a while loop
    rsdOut <- NULL #for testing while loop
    minRSD <- 100 #for testing while loop
    rtBracket = 1.5 #for testing while loop
    
    while (is.null(rsdOut) & minRSD > 30 & rtBracket < 5.1){
      #nrow(as_tibble(rsdOut)) != length(silList)) {
      
      idx_sil = silList[1]
      for(idx_sil in (silList)){
        #get experiment
        silExperiment <- master_list$templates$metaboliteNames$experiment[which(master_list$templates$metaboliteNames$newNames == idx_sil)]
        #get polarity
        silPolarity <- master_list$templates$metaboliteNames$polarity[which(master_list$templates$metaboliteNames$newNames == idx_sil)]
        #getRT
        silRT <- master_list$templates$metaboliteNames$rt[which(master_list$templates$metaboliteNames$newNames == idx_sil)]
        
        #if based on range and experiment type
        if(metaboliteExperiment == silExperiment & metabolitePolarity == silPolarity & silRT > (metaboliteRT - rtBracket) & silRT < (metaboliteRT + rtBracket)){
          
          #print(idx_sil)
          
          loopSilData <- loopDataQC[[idx_sil]]
          loopResponse <- loopMetaboliteData/loopSilData
          loopRSD <- ((sd(loopResponse, na.rm = TRUE)*100)/mean(loopResponse, na.rm = TRUE)) %>% round(2)
          
          rsdOut <- rbind(rsdOut, 
                          (c(idx_sil, loopRSD))
          )
        }
      }
      #if not matching sil then reduce requirement for experiment match
      if(is.null(rsdOut) == TRUE){
        for(idx_sil in (silList)){
          #get polarity
          silPolarity <- master_list$templates$metaboliteNames$polarity[which(master_list$templates$metaboliteNames$newNames == idx_sil)]
          #getRT
          silRT <- master_list$templates$metaboliteNames$rt[which(master_list$templates$metaboliteNames$newNames == idx_sil)]
          
          if(metabolitePolarity == silPolarity & silRT > (metaboliteRT - rtBracket) & silRT < (metaboliteRT + rtBracket)){
            
            #print(idx_sil)
            loopSilData <- loopDataQC[[idx_sil]]
            loopResponse <- loopMetaboliteData/loopSilData
            loopRSD <- ((sd(loopResponse, na.rm = TRUE)*100)/mean(loopResponse, na.rm = TRUE)) %>% round(1)
            
            rsdOut <- rbind(rsdOut, 
                            (c(idx_sil, loopRSD))
            )
          }
        }
      }
      
      #save target SIL
      if(nrow(rsdOut %>% as_tibble) > 0){minRSD <- rsdOut %>% as_tibble() %>% .$V2 %>% as.numeric() %>% min(na.rm = TRUE)}
      rtBracket <- rtBracket + 0.5
      
    }#while
    
    if(!is.null(rsdOut)){
      
      #record sil choice and rsd
      sil_choice <- (rsdOut %>% as_tibble() %>% mutate(V2 = as.numeric(V2)) %>% arrange(V2) %>% .$V1)[1]
      #rsd_value <- (rsdOut %>% as_tibble() %>% mutate(V2 = as.numeric(V2)) %>% arrange(V2) %>% .$V2)[1]
      
      #create response for imputed data
      for(idx_plate in names(master_list$data$response[[idx_dataType]])){
        master_list$data$response[[idx_dataType]][[idx_plate]] <- master_list$data$response[[idx_dataType]][[idx_plate]] %>%
          add_column(
            new_metabolite = master_list$data$peakArea[[idx_dataType]][[idx_plate]][[idx_metabolite]]/master_list$data$peakArea[[idx_dataType]][[idx_plate]][[sil_choice]]
          ) %>%
          rename_with(~idx_metabolite, "new_metabolite") 
      }
      
      
      #add sil_choice to template
      master_list$templates$metaboliteNames[[paste0("silUsed_", idx_dataType)]][which(master_list$templates$metaboliteNames$newNames == idx_metabolite)] <- 
        master_list$templates$metaboliteNames$oldNames[which(master_list$templates$metaboliteNames$newNames ==  sil_choice)]
      
      #turn matchedSilused to TRUE
      master_list$templates$metaboliteNames$matchedSilUsed[which(master_list$templates$metaboliteNames$newNames == idx_metabolite)] <- 
        FALSE
      
    } #!is.null
    
  } #idx_metabolite
  
} #for idx_dataType
  
master_list$templates$metaboliteNames

```



##RSD calculator

calculate rsd values of PQC

```{r}

rsd <- function(x) {
  if (!is.numeric(x)) stop("Input must be numeric")
  mean_x <- mean(x, na.rm = TRUE)
  sd_x <- sd(x, na.rm = TRUE)
  rsd <- (sd_x / mean_x) * 100
  return(round(rsd, 1))
}

rsdOut <- list()

idx_dataType = "imputed"
for(idx_dataType in c("imputed", "combat", "statTarget")){
  
  rsdOut[[idx_dataType]] <- list()
  
  metaboliteList = master_list$data$response[[idx_dataType]] %>% bind_rows() %>% select(!contains("sample")) %>% names()
  
  idx_metabolite = metaboliteList[1]
  
  for(idx_metabolite in metaboliteList){
    
  #get sil used for target metabolite
  sil_used = master_list$templates$metaboliteNames[[paste0("silUsed_", idx_dataType)]][which(master_list$templates$metaboliteNames$newNames == idx_metabolite)]
  
  #convert to janitor:clean_names style
  #if(!is.na(sil_used)){
     sil_used = master_list$templates$metaboliteNames$newNames[which(master_list$templates$metaboliteNames$oldNames == sil_used)]
  #}
  
  rsdOut[[idx_dataType]] <- bind_rows(
    rsdOut[[idx_dataType]],
    tibble(
      newNames = idx_metabolite,
      peakAreaRSD = master_list$data$peakArea[[idx_dataType]] %>% 
        bind_rows() %>%
        filter(sample_type == "pqc") %>%
        .[[idx_metabolite]] %>%
        rsd(.),
      silRSD = master_list$data$peakArea[[idx_dataType]] %>% 
        bind_rows() %>%
        filter(sample_type == "pqc") %>%
        .[[sil_used]] %>%
        rsd(.),
      responseRSD = master_list$data$response[[idx_dataType]] %>% 
        bind_rows() %>%
        filter(sample_type == "pqc") %>%
        .[[idx_metabolite]] %>%
        rsd(.)
    )
  )
  
  } #idx_metabolite

}#idx_dataType

master_list$summary_tables[["rsdOut"]] <-  rsdOut

```


#.
# review SIL internal standard match
## export template

*USER INPUT REQUIRED*

opportunity to swap out some internal standards

```{r}

if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/internalStandardSelection/"))){
  dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/internalStandardSelection/"))
}


 write_csv(
   master_list$templates$metaboliteNames %>%
     select(-silUsed_imputed, -silUsed_combat, -silUsed) %>%
     left_join(
       .,
       master_list$summary_tables$rsdOut$statTarget, 
       by = "newNames"
     ) %>% 
     filter(!grepl("sil", newNames)) %>%
     arrange(desc(matchedSilUsed), oldNames),
   file = paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/internalStandardSelection/",
                 Sys.Date(), "_internStandards_forReview.csv")
 )


```

*USER INPUT REQUIRED*

edit and read in template

```{r}

#make any edits to the silUsed column and then change file name to edited

#make sure any names for IS use the tidyName column

updated_metaboliteNames <- read_csv(file = paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/internalStandardSelection/",
                 Sys.Date(), "_internStandards_edited.csv")
                 )

#update master list

master_list$templates$metaboliteNames <- master_list$templates$metaboliteNames %>%
  select(-silUsed_statTarget, -silUsed) %>%
  left_join(
    by = "tidyName",
    updated_metaboliteNames %>% 
      select(tidyName, silUsed_statTarget)
  ) 


```


# control charts

## run order plots

```{r}

if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/controlCharts/"))){
  dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/controlCharts/"))
}

#add a vairable that adapts page width for number of plates

pageWidth <- length(master_list$project_details$plateList)

 #open pdf device
  pdf(
    onefile = TRUE, width = (15 + pageWidth), height = 10,
    file = paste0(
                  master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/controlCharts/", Sys.Date(), "_", master_list$project_details$project_name, "_SPE_", master_list$project_details$wash, "_all_metabolites_controlCharts.pdf"
                
    )
  )
  
  
 
  
idx_metabolite = "x3_haa"
for(idx_metabolite in (master_list$data$response$statTarget %>% bind_rows() %>% select(!contains("sample")) %>% names())){
  
  
  # --------------------------------------------------------------
  ## Data preparatoion
  # -------------------------------------------------------------- 
  
  ### for peak area
  
  #extract SIL to plot
  plotSIL <- master_list$templates$metaboliteNames$silUsed_statTarget[which(master_list$templates$metaboliteNames$newNames == idx_metabolite)]
  #get janitor::tidyName
  plotSIL <- master_list$templates$metaboliteNames$newNames[which(master_list$templates$metaboliteNames$oldNames == plotSIL)]
  
  # get peak areas to plot
  plotArea <- master_list$data$peakArea$imputed %>% bind_rows() %>% 
    select(sample_run_index, sample_name, sample_type, sample_plate, all_of(idx_metabolite), all_of(plotSIL)) %>%
    setNames(c("sample_run_index", "sample_name", "sample_type", "sample_plate", "area", "silArea")) %>%
    left_join(
      .,
      master_list$data$response$impute %>% bind_rows() %>% 
        select(sample_name, all_of(idx_metabolite)) %>% setNames(c("sample_name", "response")),
      by = "sample_name"
    )
  
  
  #combine into plotArea
  plotData <- plotArea %>%
    filter(sample_type != "vltr") %>%
    pivot_longer(
      cols = c("area", 
               "silArea",  
               "response"),
      names_to = "dataType",
      values_to = "value"
    ) 
  
  pqcData = plotData %>%
    filter(sample_type == "pqc")
  
  #get median values
  loopMedian =  pqcData %>% 
    group_by(dataType) %>% 
    summarize(median_val = median(value, na.rm = TRUE))
  
  #loop mean
  loopMean = pqcData %>% 
    group_by(dataType) %>% 
    summarize(mean_val = mean(value,  na.rm = TRUE))
  
  #loop sd
  loopSd = pqcData %>% 
    group_by(dataType) %>% 
    summarize(sd_val = sd(value,  na.rm = TRUE))
  
  #loopRsd
   loopRsd = loopSd %>%
     select(dataType) %>%
     add_column(
       rsd_val = (loopSd$sd_val / loopMean$mean_val) *100
     )
   
   
   # get peak areas to plot
  plotAreaCor <- master_list$data$peakArea$statTarget %>% bind_rows() %>% 
    select(sample_run_index, sample_name, sample_type, sample_plate, all_of(idx_metabolite), all_of(plotSIL)) %>%
    setNames(c("sample_run_index", "sample_name", "sample_type", "sample_plate", "area", "silArea")) %>%
    left_join(
      by = "sample_name",
      .,
      master_list$data$response$statTarget %>% bind_rows() %>% 
        select(sample_name, all_of(idx_metabolite)) %>% setNames(c("sample_name", "response"))
    )
  
  #combine into plotArea
  plotDataCor <- plotAreaCor %>%
    filter(sample_type != "vltr") %>%
    rename(areaCor = area, silAreaCor = silArea, responseCor = response) %>%
    pivot_longer(
      cols = c("areaCor", 
               "silAreaCor",  
               "responseCor"),
      names_to = "dataType",
      values_to = "value"
    )  %>%
    mutate(sample_type = paste0(sample_type, "Cor")) 
  
  pqcDataCor = plotDataCor %>%
    filter(sample_type == "pqcCor")
  
  #get median values
  loopMedianCor =  pqcDataCor %>% 
    group_by(dataType) %>% 
    summarize(median_val = median(value, na.rm = TRUE))
  
  #loop mean
  loopMeanCor = pqcDataCor %>% 
    group_by(dataType) %>% 
    summarize(mean_val = mean(value,  na.rm = TRUE))
  
  #loop sd
  loopSdCor = pqcDataCor %>% 
    group_by(dataType) %>% 
    summarize(sd_val = sd(value,  na.rm = TRUE))
  
  #loopRsd
   loopRsdCor = loopSdCor %>%
     select(dataType) %>%
     add_column(
       rsd_val = (loopSdCor$sd_val / loopMeanCor$mean_val) *100
     )
   
   #join data
   plotDataALL <- bind_rows(
     plotData,
     plotDataCor
   ) %>%
   mutate(sample_type = factor(sample_type, levels = c("sample", "sampleCor", "pqc", "pqcCor", "ltr", "ltrCor", "NA"), ordered = TRUE)) %>%
     arrange(sample_name, dataType)
   
    
  ####--------------------
   #add facets
  ####--------------------
   
  
    
  #add facet column
  plotDataALL[["facet"]] <- NA
   
   #area
   plotDataALL$facet[which(plotDataALL$dataType == "area")] <- paste0("1. area;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "area")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "areaCor")] %>% floor(), "%")
   
   #areaCor
   plotDataALL$facet[which(plotDataALL$dataType == "areaCor")] <- paste0("1. area;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "area")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "areaCor")] %>% floor(), "%")
   
   #silArea
   plotDataALL$facet[which(plotDataALL$dataType == "silArea")] <- paste0("2. silArea;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "silArea")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "silAreaCor")] %>% floor(), "%")
   
    #silAreaCor
   plotDataALL$facet[which(plotDataALL$dataType == "silAreaCor")] <- paste0("2. silArea;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "silArea")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "silAreaCor")] %>% floor(), "%")
                    
   #response                                                                                                               
  plotDataALL$facet[which(plotDataALL$dataType == "response")] <- paste0("3. response;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "response")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "responseCor")] %>% floor(), "%")
  
  #responseCor
    plotDataALL$facet[which(plotDataALL$dataType == "responseCor")] <- paste0("3. response;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "response")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "responseCor")] %>% floor(), "%")
  
  
  #add facet for median PQC hline 
  loopMedian[["facet"]] <- NA
  #area
  loopMedian$facet[which(loopMedian$dataType == "area")] <- paste0("1. area;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "area")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "areaCor")] %>% floor(), "%")
  #silArea
  loopMedian$facet[which(loopMedian$dataType == "silArea")] <- paste0("2. silArea;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "silArea")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "silAreaCor")] %>% floor(), "%")
  #response
  loopMedian$facet[which(loopMedian$dataType == "response")] <- paste0("3. response;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "response")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "responseCor")] %>% floor(), "%")
  
   #add facet for median PQC hline 
  loopMedianCor[["facet"]] <- NA
  #area
  loopMedianCor$facet[which(loopMedianCor$dataType == "areaCor")] <- paste0("1. area;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "area")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "areaCor")] %>% floor(), "%")
  #silArea
  loopMedianCor$facet[which(loopMedianCor$dataType == "silAreaCor")] <- paste0("2. silArea;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "silArea")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "silAreaCor")] %>% floor(), "%")
  #response
  loopMedianCor$facet[which(loopMedianCor$dataType == "responseCor")] <- paste0("3. response;\n", "PQC %RSD = ", loopRsd$rsd_val[which(loopRsd$dataType == "response")] %>% floor(), "%;\n corrected PQC %RSD = ", loopRsdCor$rsd_val[which(loopRsdCor$dataType == "responseCor")] %>% floor(), "%")
  
  
  
  

  
# --------------------------------------------------------------
## PLOT
# --------------------------------------------------------------  
  
  #set plot title
  plotTitle <- paste0(
    master_list$templates$metaboliteNames$tidyName[which(master_list$templates$metaboliteNames$newNames == idx_metabolite)],
    "; SIL used =  ",
    master_list$templates$metaboliteNames$tidyName[which(master_list$templates$metaboliteNames$newNames == plotSIL)],
    "; matched SIL = ", 
    master_list$templates$metaboliteNames$matchedSilUsed[which(master_list$templates$metaboliteNames$newNames == idx_metabolite)]
  )
  
  #create plot
  ccPlot <- ggplot(data = plotDataALL, aes(x = sample_run_index, y = value)) +
    geom_point(data = plotDataALL %>% filter(grepl("sample", sample_type)), aes(colour = sample_type,fill = sample_type, shape = sample_type, size = sample_type)) +
    geom_point(data = plotDataALL %>% filter(!grepl("sample", sample_type)), aes(colour = sample_type,fill = sample_type, shape = sample_type, size = sample_type)) +
    theme_bw() +
    scale_shape_manual(values = c("sample" = 0, "sampleCor" = 23, "pqc" = 22, "pqcCor" = 23, "ltr" = 22, "ltrCor" = 23)) +
    scale_size_manual(values = c("sample" = 2, "sampleCor" = 2, "pqc" = 3, "pqcCor" = 3, "ltr" = 3, "ltrCor" = 3)) +
    scale_color_manual(values = c("sample" = "grey", "sampleCor" = "grey", "pqc" = "orchid", "pqcCor" = "black", "ltr" = "seagreen3", "ltrCor" = "black")) +
    scale_fill_manual(values = c("sample" = "grey", "sampleCor" = "grey", "pqc" = NA, "pqcCor" = "orchid", "ltr" = NA, "ltrCor" = "seagreen3")) +
    geom_hline(data= loopMedian, aes(yintercept=median_val), colour = "orchid4", linetype ="dotted", linewidth = 1) +
    geom_hline(data= loopMedianCor, aes(yintercept=median_val), colour = "orchid4", linetype ="solid", linewidth = .8) +
    facet_wrap(facets = "facet", ncol = 1, scales = "free_y") +
    scale_y_continuous(labels = scales::label_scientific()) +
    ggtitle(plotTitle) +
    theme(legend.position = "none")
    
  
  
  ####------------------------
  ## VIolin plots
  ####------------------------
  
  violinData = plotDataALL %>% 
    #filter(dataType == "responseBatchCorrected" | dataType == "response") %>%
    mutate(sample_plate = sub("a", "", sample_plate)) %>%
    mutate(sample_plate = sub("b", "", sample_plate)) %>%
    #add columns tio violinData
    add_column(
      side = ifelse(grepl("Cor", .$sample_type), "r", "l"),
      sample_type_x = gsub("Cor", "", .$sample_type)
    ) 
  
  #add max/min dummy data to normalise axis on facet_wrap
  for(idx_dataType in unique(violinData$dataType)){
  #work around to normalise axis on per plate plot
  maxValue <- violinData %>%
    filter(dataType == idx_dataType) %>%
    .$value %>%
    max()
  minValue <- violinData %>%
    filter(dataType == idx_dataType) %>%
    .$value %>%
    min()
  
  #make dummy data per plate for plot control
  dummyData = list()
  for(idx_plate in unique(violinData$sample_plate)){
  dummyData <- dummyData %>%
    bind_rows(.,
              bind_rows(  
                (violinData %>%
                   filter(dataType == idx_dataType) %>%
                   filter(value == maxValue))[1,],
                (violinData %>%
                   filter(dataType == idx_dataType) %>%
                   filter(value == minValue))[1,]
              ) %>% mutate(
                sample_plate = idx_plate,
                sample_run_index = NA) 
    ) 
  } # idx plate\
  
  
  for(idx_sampleType in unique(violinData$sample_type_x)){
    dummyData <- dummyData %>%
      bind_rows(.,
                bind_rows(  
                  (violinData %>%
                     filter(dataType == idx_dataType) %>%
                     filter(value == maxValue))[1,],
                  (violinData %>%
                     filter(dataType == idx_dataType) %>%
                     filter(value == minValue))[1,]
                ) %>% mutate(
                  sample_type_x = idx_sampleType,
                  sample_run_index = NA) 
      ) 
  } # idx sampleType
  
  #add dummy data to violinData
  violinData <- violinData %>%
    bind_rows(.,
              dummyData)
  }
  
  #change sample type for colour control
  violinData$sample_type[is.na(violinData$sample_run_index)] <- "NA"
  
  #mutate to factor
  violinData <- violinData %>%
    mutate(
      sample_type_x = factor(sample_type_x, levels = c("pqc", "ltr", "sample", "NA"), ordered = TRUE),
      sample_plate_idx = factor(sample_plate)
    ) %>%
    mutate(
      sample_plate_idx = as.integer(sample_plate_idx)
    ) %>%
    mutate(
      sample_plate_idx = as.factor(sample_plate_idx)
    )
    

#make plot
violinPlot <- ggplot(data = violinData, aes(x = sample_type_x, y = value, fill = sample_type, color = sample_type)) +
  theme_bw() +
  # Split violin using gghalves
  geom_half_violin(
    data = violinData %>% filter(grepl("Cor", sample_type)),
    side = "r",
    position = position_nudge(x = 0), 
    trim = TRUE, alpha = 1, scale = "width"
  ) +
  geom_half_violin(
    data = violinData %>% filter(!grepl("Cor", sample_type)),
    side = "l",
    position = position_nudge(x = 0), 
    trim = TRUE, alpha = 0.25, scale = "width"
  ) +
  geom_hline(data= loopMedian, aes(yintercept=median_val), colour = "orchid1", linetype ="dotted", linewidth = 1) +
  geom_hline(data= loopMedianCor, aes(yintercept=median_val), colour = "orchid4", linetype ="solid", linewidth = .8) +
  geom_half_boxplot(
    data = violinData %>% filter(grepl("Cor", sample_type)),
    side = "r",
    position = position_nudge(x = 0.1), 
    notch = FALSE, alpha = 1, outlier.shape = NA,
    color = "grey30", fill = NA, center = TRUE, width = 0.25
  ) +
  geom_half_boxplot(
    data = violinData %>% filter(!grepl("Cor", sample_type)),
    side = "l",
    position = position_nudge(x = -0.1), 
    notch = FALSE, alpha = 1, outlier.shape = NA,
    color = "grey30", fill = NA, center = TRUE, width = 0.25
  ) +
  facet_wrap(facets = c("facet"), ncol=1, scales="free_y") +
  scale_fill_manual(values = c("sample" = "grey", "sampleCor" = "grey", "pqc" = "orchid", "pqcCor" = "orchid", "ltr" = "seagreen3", "ltrCor" = "seagreen3", "NA"= NA)) +
  scale_color_manual(values = c("sample" = "grey", "sampleCor" = "grey", "pqc" = "orchid", "pqcCor" = "orchid", "ltr" = "seagreen3", "ltrCor" = "seagreen3", "NA"= NA)) +
  scale_y_continuous(labels = scales::label_scientific()) +
  #scale_shape_manual(values = c("sample" = 0, "sampleCor" = 23, "pqc" = 22, "pqcCor" = 23, "ltr" = 22, "ltrCor" = 23)) +
  ggtitle("all plates") +
  theme(legend.position = "none")





  #make a plot per plate
  violinPlotplate <- ggplot(data = violinData, aes(x = sample_plate_idx, y = value, fill = sample_type, color = sample_type)) +
  theme_bw() +
  # Split violin using gghalves
  geom_half_violin(
    data = violinData %>% filter(grepl("Cor", sample_type)),
    side = "r",
    position = position_nudge(x = 0), 
    trim = TRUE, alpha = 1, scale = "width"
  ) +
  geom_half_violin(
    data = violinData %>% filter(!grepl("Cor", sample_type)),
    side = "l",
    position = position_nudge(x = 0), 
    trim = TRUE, alpha = 0.25, scale = "width"
  ) +
  geom_hline(data= loopMedian, aes(yintercept=median_val), colour = "orchid1", linetype ="dotted", linewidth = 1, alpha = 1) +
  geom_hline(data= loopMedianCor, aes(yintercept=median_val), colour = "orchid4", linetype ="solid", linewidth = .8, alpha = 1) +
  geom_half_boxplot(
    data = violinData %>% filter(grepl("Cor", sample_type)),
    side = "r",
    position = position_nudge(x = 0.1), 
    notch = FALSE, alpha = 1, outlier.shape = NA,
    color = "grey30", fill = NA, center = TRUE, width = 0.25
  ) +
  geom_half_boxplot(
    data = violinData %>% filter(!grepl("Cor", sample_type)),
    side = "l",
    position = position_nudge(x = -0.1), 
    notch = FALSE, alpha = 1, outlier.shape = NA,
    color = "grey30", fill = NA, center = TRUE, width = 0.25
  ) +
  facet_wrap(facets = c("facet", "sample_type_x"), ncol = length(unique(violinData$sample_type_x)), scales = "free_y") +
  scale_fill_manual(values = c("sample" = "grey", "sampleCor" = "grey", "pqc" = "orchid", "pqcCor" = "orchid", "ltr" = "seagreen3", "ltrCor" = "seagreen3", "NA"=NA)) +
  scale_color_manual(values = c("sample" = "grey", "sampleCor" = "grey", "pqc" = "orchid", "pqcCor" = "orchid", "ltr" = "seagreen3", "ltrCor" = "seagreen3", "NA"=NA)) +
  scale_y_continuous(labels = scales::label_scientific()) +
  #scale_shape_manual(values = c("sample" = 0, "sampleCor" = 23, "pqc" = 22, "pqcCor" = 23, "ltr" = 22, "ltrCor" = 23)) +
  ggtitle("data split per plate") +
  theme(legend.position = "none")
    
  
  
  combinedPlot <- cowplot::plot_grid(
    ncol = 3,
    rel_widths = c(0.35,0.15,(0.05 * pageWidth)),
    ccPlot,
    violinPlot,# NULL, nrow = 1, rel_widths = c(0.2,0.6,0.2)),
    violinPlotplate
    )
  

  print(combinedPlot)
    
  }

dev.off()


```

# pca

```{r}

rsdPerformance <- master_list$summary_tables$rsdOut$statTarget

X= master_list$data$response$statTarget %>% 
  bind_rows() %>%
  filter(sample_type %in% c("sample", "ltr", "pqc")) %>%
  select(
    all_of(
      (rsdPerformance %>% filter(responseRSD<30) %>% .$newNames)
    )
  )

Y= master_list$data$response$statTarget %>% 
  bind_rows() %>% 
  filter(sample_type %in% c("sample", "ltr", "pqc")) %>%
  .$sample_type

plate = master_list$data$response$statTarget %>% 
  bind_rows() %>% 
  filter(sample_type %in% c("sample", "ltr", "pqc")) %>%
  .$sample_plate %>%
  gsub("a", "", .) %>%
  gsub("b", "", .)

pcaMod <- ropls::opls(
  x = X+1,
  predI = 3, 
  log10L = TRUE,
  scaleC = 'pareto',
  fig.pdfC = 'none', 
  algoC = 'default'
  )


plotDataALL = bind_cols(
  pcaMod@scoreMN %>% as_tibble(), 
  Y %>% as_tibble() %>% rename("sample_type" = value),
  plate %>% as_tibble() %>% rename("sample_plate" = value)
)


plot.PCA <- ggplot(data = plotDataALL, aes(x=p1, y=p2, fill = sample_type, shape = sample_type, color = sample_type)) +
  geom_point(data = plotDataALL %>% filter(sample_type == "sample"), size = 1.5) +
  geom_point(data = plotDataALL %>% filter(sample_type == "ltr"), size = 2.5) +
  geom_point(data = plotDataALL %>% filter(sample_type == "pqc"), size = 2.5) +
  scale_shape_manual(values = c("sample" = 21, "pqc" = 23, "ltr" = 21)) +
  scale_color_manual(values = c("sample" = "black", "pqc" = "black", "ltr" = "black")) +
  scale_fill_manual(values = c("sample" = "white", "pqc" = "orchid", "ltr" = "seagreen3")) +
  theme_bw() +
  ggtitle(
    paste0(
      "PCA on batch corrected response data;\n(samples = ", 
      nrow(master_list$data$response$combat %>% bind_rows() %>% filter(sample_type == "sample")),
      "; PQC = ",
      nrow(master_list$data$response$combat %>% bind_rows() %>% filter(sample_type == "pqc")),
      "; LTR = ",
      nrow(master_list$data$response$combat %>% bind_rows() %>% filter(sample_type == "ltr")),
      ";\nmetabolites = ",
      ncol(X),
      ")"
    )
    );plot.PCA

plot.PCA.Plate <- ggplot(data = plotDataALL, aes(x=p1, y=p2, shape = sample_type, color = sample_type, fill = sample_plate)) +
  geom_point(data = plotDataALL %>% filter(sample_type == "sample"), size = 2) +
  geom_point(data = plotDataALL %>% filter(sample_type == "ltr"), size = 2.5) +
  geom_point(data = plotDataALL %>% filter(sample_type == "pqc"), size = 2.5) +
  scale_shape_manual(values = c("sample" = 22, "pqc" = 24, "ltr" = 25)) +
  scale_color_manual(values = c("sample" = "grey", "pqc" = "black", "ltr" = "black")) +
  #scale_fill_manual(values = c("RARp003" = "white", "RARp004" = "orchid", "RARp005" = "seagreen3")) +
  theme_bw() +
  guides(fill = guide_legend(override.aes = list(shape=21))) +
  ggtitle(
    paste0(
      "PCA on batch corrected response data;\n(samples = ", 
      nrow(master_list$data$response$combat %>% bind_rows() %>% filter(sample_type == "sample")),
      "; PQC = ",
      nrow(master_list$data$response$combat %>% bind_rows() %>% filter(sample_type == "pqc")),
      "; LTR = ",
      nrow(master_list$data$response$combat %>% bind_rows() %>% filter(sample_type == "ltr")),
      ";\nmetabolites = ",
      ncol(X),
      ")"
    )
  );plot.PCA.Plate





  #save pca
if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/pca/"))){
  dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/pca/"))
}

 #open pdf device
  pdf(
    onefile = TRUE, width = 15, height = 10,
    file = paste0(
                  master_list$project_details$project_dir, "/", master_list$project_details$wash, "/plots/pca/", Sys.Date(), "_", master_list$project_details$project_name, "_pca_filteredMetabolites.pdf"
    )
  )
  
  print(
    cowplot::plot_grid(
      plot.PCA, 
      plot.PCA.Plate
    )
  )
  
  
  
  dev.off()
  
```

# exportData

```{r}
  
#export csv
  
  exportData <- master_list$data$response$statTarget %>% bind_rows() %>%
    select(
      contains("sample"),
      all_of((master_list$summary_tables[["rsdOut"]]$statTarget %>% filter(responseRSD<30) %>% arrange(newNames) %>% .$newNames))
    )
      
for(idx_metabolite in exportData %>% select(!contains("sample")) %>% names()){
  
  name_from <- idx_metabolite
  name_to <- master_list$templates$metaboliteNames$tidyName[which(master_list$templates$metaboliteNames$newNames==idx_metabolite)]
  
  exportData <- exportData %>%
    rename_with(~name_to, all_of(name_from))
  
}
  
  if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/csvExport/"))){
    dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/csvExport/"))
  }
  
  write_csv(exportData,
            file =  paste0(
                  master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/csvExport/", Sys.Date(), "_", 
                  master_list$project_details$user_name, "_",
                  master_list$project_details$project_name, "_", 
                  master_list$project_details$spe, "_",
                  master_list$project_details$wash, "_metaboliteDATA.csv"
                )
              )
```

#.

# checkpoint save

#.


```{r, eval=TRUE, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}

if(!dir.exists(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/rda/"))){
  dir.create(paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/rda/"))
}

save(
  master_list,
  file = paste0(master_list$project_details$project_dir, "/", master_list$project_details$wash, "/data/rda/",
                Sys.Date(), "_",
                master_list$project_details$project_name,
                "_SPE_AB_checkpoint_2.rda"
  )
)

# load(file = "C:\\Users\\300828F\\Murdoch University\\Australian National Phenome Centre - Technical & Operational\\Laboratory QA QC\\Control Chart Reviews\\SPE-MCX\\lgwTest\\AB\\data\\rda\\2025-07-25_SPRESTO_SPE_AB_checkpoint_1.rda")

```

